---
title: "Sankey Plots"
output: html_notebook
---

```{r}
# 2020-10-26_07:35 
save.image("../R/2020-10-26_sankey_plots.RData")

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read data
  • Copied from:`../aparna/homefiles/Fall2019_finalclustering_secondtrimester/_neocortex_*_byarea_clustermarkers.txt`
```{r read-gene-markers}
dir <- "../../../area_specific_markers/tbls_markers_by_celltype_and_area/"

files <- fs::dir_ls(dir, regexp = "txt")
       
cell.type <- c("rg", "neuron", "ipc") %>% as.list %>% set_names(unlist(.))

# files <- files[str_detect(tolower(files), cell.type$rg)]

read_and_add <- function(file) {
  
     metadata <- str_match(file, ".*/(.*)_neocortex_([:alnum:])_byarea.*") %>% tolower
     df = read.table(file)
     
     df$individual = metadata[2]
     df$celltype = metadata[3]
     df$file = file
     
     df %<>% select(gene, everything()) %>% rename(area = "cluster")
     return(df)
}


df.markers <- map_dfr(.x = files,
                   .f = read_and_add)

df.markers <- df.markers %>% mutate( pct.2 = (pct.2 + 0.001),
                                     gene.ratio = (pct.1 / pct.2),
                                     gene.score = avg_logFC * gene.ratio,
                                    
                                    across(.cols = where(is.numeric), .fns = ~round(.x, 5)),
                                    across(.cols = c(where(is.character), -gene), .fns = ~tolower(.x))
)

df.markers %>% plot_bar()

df.markers[complete.cases(df.markers), ] %>% select(-gene, -file) %>% inspect_cat %>% show_plot()
# Outliers?
markers[!complete.cases(markers), ] %>% plot_bar
markers[!complete.cases(markers), ] %>% inspect_cat() %>% show_plot()

# Remove outlier cells
df.markers <- df.markers[complete.cases(df.markers), ]
```

```{r plot-counts-bar}

.stash <- new_environment()
.stash$df.markers <- df.markers

df.markers <- df.markers %>% 
                 filter(celltype %in% c("rg", "ipc", "neuron")) %>%
                 left_join(colors.area) %>% 
                 mutate(area = factor(area, levels = colors.area$area),
                        region = factor(region, levels = colors.area$area),
                        celltype = factor(celltype, levels = c("rg", "ipc", "neuron"))
                        )


  # arrange(match(area, colors.area))
df.markers %>% ggplot +
  geom_bar(aes(x = area, fill = area)) + 
  facet_grid(rows = vars(celltype), 
             cols = vars(individual)) +
  scale_fill_manual(values =  colors.area$color) +
  theme_minimal()

df.markers %>% inspect_cat() %>% show_plot()


df.markers[!complete.cases(df.markers), ]





colors.area %<>% mutate(category = str_replace(category, "somato", "somatosensory")) %>%
  rename(area = "category")
```


```{r plot-counts-smooth}
df.markers %>% group_by(celltype) %>% count(individual, area) %>%
    ggplot(aes(x = individual, y = n, group = area, color = area)) + 
  geom_point(alpha = 0.3) +
  geom_smooth(method = "loess", se = FALSE, alpha = 0.5) +
facet_grid(rows = vars(celltype), cols = vars(area), scale = "free_y")  + 
  scale_color_manual(values =  colors.area$color) +
  theme_minimal()
```

# fxn: buildSankey

```{r}
# ARGUMENTS
# ===========
# 1. A dataframe `markers`
# 2. An optional stage (early/mid/late).
# 3. An out.dir

# OUTPUT
  pairs
  unique.genes

# TODO make the grouping variable be argument

# What does this function do?

# markers.plot.df <- df.markers.distinct

buildSankey <- function(markers.plot.df = df.markers, 
                        stage = NULL, 
                        dir = "") {
  
      plot.name <- first(markers.plot.df$celltype)

      if (! is.null(stage)) {
        
         markers.plot.df  <- markers.plot.df  %>% dplyr::filter(str_detect(region.stage, stage)) %>%
         
         plot.name <- paste0(plot.name, "_", stage)
         
      }
      
      #############################################################################################
      
      markers.plot.df %>% arrange(group, stage, region)
      
      
      # Calculate pairwise intersection of genes.
      pairs.raw <-  widyr::pairwise_count(markers.plot.df, 
                                      item = group, 
                                      feature = gene, 
                                      sort = TRUE, diag = TRUE, upper = TRUE) %>% 
                                      arrange(item1, item2)

      # For RG -> neuron 
      # pairs <- pairs.raw %>% filter(item1 %>% str_detect("rg") & 
      #                              item2 %>% str_detect("neuron")
      #                          )
      
      
      # UNIQUE GENES 
      # . . . . . . . 
      # Get unique genes for each group (not shared with any other [celltype]:region:stage group)
      
      unique.genes <- lapply(unique(markers.plot.df$group), 
                             
                             function(x) {

                                 outgroup <- markers.plot.df %>% 
                                               dplyr::filter(! group == x)
                                 group <- markers.plot.df %>% 
                                               dplyr::filter(group == x)
                
                                  setdiff(group$gene, outgroup$gene)
                             }
                      ) %>% set_names(unique(markers.plot.df$group))
      
      # Get counts of unique genes for each [celltype]:region:stage group
      unique.genes <- unique.genes %>% tibble(item1 = names(.),
                                              item2 = paste0(item1, "_unique"),
                                              n = lapply(., length) %>% unlist,
                                              genes = .
                                              
                              )

      genes.intersect <- purrr::map2(.x = pairs.raw$item1, .y = pairs.raw$item2, 
                      .f = function(.x, .y) {
                              a <- df.markers.distinct %>% filter(group == .x)
                              b <- df.markers.distinct %>% filter(group == .y)
                              intersect(a$gene, b$gene)
                      }
                                  ) %>% tibble(genes = .)
      
      pairs <- pairs.raw %>% bind_cols(genes.intersect) %>%
            bind_rows(unique.genes)
      
      
         # Get # of shared genes
         # {r 20201021_1604}

      pairs_shared <- pairs %>% filter(item2 %>% str_detect("unique") | item1 == item2) %>% 
            group_by(item1) %>% summarise(n = first(n) - n[2]) %>% 
                add_column(item2 = paste0(.$item1, "_shared"), .after = 1) 

      pairs <- pairs %>%
            bind_rows(pairs_shared)  %>%
             mutate(item2 = case_when(item1 == item2 ~ paste0(item2, "_all"),
                                      TRUE ~ item2 )) %>% 
               arrange(item1, desc(item2))
        
        # 1,289 comparisons 
      
            # To see the numbers of genes shared, unique and all for each group:
               # pairs %>% filter(item2 %in% c("all", "shared", "unique")) %>% arrange(item1)
      

      ##############################################################################################

      all.to.unique <- pairs %>% dplyr::filter(item1 == item2) %>%
        mutate(n = unique.genes$counts[item1,],
               item1 = paste0("all_", item1),
               item2 = paste0("unique_", item2)
              )

      all.to.shared <- pairs %>% dplyr::filter(item1 == item2) %>%
        mutate(n = n - unique.genes$counts[item1, ],
               item1 = paste0("all_", item1),
               item2 = paste0("shared_", item2)
              )

      unique <- pairs %>% dplyr::filter(item1 == item2) %>%
        mutate(n = 0,
               item1 = paste0("unique_", item1),
               item2 = paste0("to_", item2)
              )

      shared <- pairs %>% dplyr::filter(!item1 == item2) %>%
                   mutate(item1 = paste0("shared_", item1),
                          item2 = paste0("to_", item2)
                          )

      pairs <- rbind(all.to.unique, all.to.shared, unique, shared)
      
   



      # Indicate order of nodes on graph
      # . . . . . . . . . . . . . . . . 
      
      levels.sankey <- markers.plot.df$group %>% unique
        
        # lapply(c("pfc", "msp", "temporal", "v1"), 
        #                      function(x) 
        #                        paste(x, c("early", "mid", "late"), sep = "_")
        #                      ) %>% unlist

      levels.sankey <- lapply(levels.sankey, 
                              FUN = function(x) paste0(c("all_", "unique_", "shared_", "to_"), x)
                              ) %>% 
                          unlist %>%
                          data.frame(name = .) 

      levels.sankey.x <- levels.sankey %>% dplyr::filter(name %>% str_detect("all_"))
      levels.sankey.y <- levels.sankey %>% dplyr::filter(name %>% str_detect("unique_|shared_"))
      levels.sankey.z <- levels.sankey %>%  dplyr::filter(name %>% str_detect("to_"))
      
      levels.sankey <- rbind(levels.sankey.x, levels.sankey.y, levels.sankey.z) 
      
      # . . . . . . 
      # Create nodes and links dfs for Sankey 
      # . . . . . . 
      
      nodes <- levels.sankey %>% dplyr::filter(name %in% pairs$item1 | name %in% pairs$item2 ) %>% 
                               mutate(ID = (as_factor(name) %>% as.integer) - 1,
                                      node.group = name %>% str_extract(pattern = str_c(colors.area$category, 
                                                                                        collapse = "|")))

       links <- inner_join(x = nodes[ , c("name", "ID")], y = pairs, by = (c("name" = "item1"))) %>%
                  inner_join(x = ., y = nodes[c("name", "ID")], 
                             by = (c("item2" = "name"))
                             ) %>% 
                    # select(source, item1, target, item2, n) %>%
                    set_names(c("source.name", "source", "target.name", "n", "target")) %>% 
                      mutate(link.group.source = str_extract(source.name,
                                                             pattern = str_c(colors.area$category, 
                                                               collapse = "|")
                                                             ),
                              link.group.target = str_extract(target.name,
                                                              pattern = str_c(colors.area$category, 
                                                              collapse = "|")
                                                    )
                                           ) %>% as.tibble %>% dplyr::filter(n >= 3)
       
           #       arrange(source, target)   %>%
              

      
      # Colors
      colors <- c("pfc" = "#F94171", "msp" = "#FFA750", "temporal" = "#3182BD", "v1" = "#9E9AC8")

     # nodes.a <- links %>% distinct(item1, source) %>% set_names(c("name", "ID"))
    #  nodes.b <- links %>% distinct(item2, target) %>% set_names(c("name", "ID"))
      
  #  nodes <- rbind(nodes.a, nodes.b) %>% 
                 
      
 #     nodes <- pairs %>% select(item1, item2) %>% unlist(c(.$item1, .$item2), use.names = FALSE) %>% unique %>%
 #        factor(levels = levels.sankey$name) %>% data.frame(name = .) 

      links.x <- links %>% filter(item1 %>% str_detect("all_rg|shared_rg"))
                                
    # sankey <- 
      
  x <-  sankeyD3::sankeyNetwork(# LinkGroup = "color",
                                linkType = "l-bezier",
                                curvature = 0.7 ,
                                linkColor = "color",
                                highlightChildLinks = TRUE, 
                                        dragX = TRUE, dragY = TRUE, 
                                        # linkOpacity = 1, 
                                        # doubleclickTogglesChildren = TRUE, 
                                           
                                        linkGradient  = TRUE, 
                              
                              zoom = TRUE,
                                        nodeCornerRadius = 5,
                                        nodePadding = 20,
                              # margin = list(top = 100, bottom = 100),
                              # sinksRight = FALSE,
                              # height = 250,
                              # width = 500,
                    Links = df.links[ , c("source", "target", "n", "color")],
                    Nodes = rbind(df.links[ ,c("source.name", "source", "color.rg")] %>% set_names(c("name", "ID", "color")),
                                  df.links[ ,c("target.name", "target", "color")] %>% set_names(c("name", "ID", "color"))
                                  ) %>% distinct, 
                    Source = "source",
                    Target = "target", 
                    Value = "n", 
                    NodeID = "name",
                    NodeColor = "color",
                    units = "genes", # optional units name for popups
                    # NodeGroup = "node.group",
                    # LinkGroup = "link.group",
                    fontSize = 12, 
                    nodeWidth = 30,
                    iterations = 0,
                    colourScale = 'd3.scaleOrdinal().range(["#F94171", "#FFA750", "#9E9AC8", "#3182BD"])'
                   )
     
  x %>% listviewer::jsonedit()

      # sankey <- htmlwidgets::prependContent(sankey, htmltools::tags$h1(plot.name))
      # file.html <- paste0(dir, "sankey_" , plot.name, ".html")
      # htmlwidgets::saveWidget(widget = sankey, file = file.html, title = plot.name, selfcontained = TRUE)

      return(list(sankey.plot = sankey, 
                  links = links, 
                  nodes = nodes, 
                  pairs = pairs, 
                  pairs.raw = pairs.raw,
                  levels = levels.sankey, 
                  unique.genes = unique.genes, 
                  markers = markers.plot.df))
}

htmlwidgets::saveWidget(x, file = "neurons_RG.html")
```


```{r}
# https://stackoverflow.com/questions/24614391/intersect-all-possible-combinations-of-list-elements

# Get the combinations of names of list elements
nms <- combn( unique(df.markers.distinct$group) , 2 , FUN = paste0 , collapse = "" , simplify = FALSE)

# Make the combinations of list elements
ll <- combn( df.msrkers.   , 2 , simplify = FALSE )

# Intersect the list elements
out <- lapply( ll , function(x) length( intersect( x[[1]] , x[[2]] ) ) )

# Output with names
setNames( out , nms )
```

```{r getGeneLists}
# [x] this goes in `buildSankey`
  # 2020-10-23 Moved up 

# TODO
# Proportion of genes unique vs shared? 
```


```{r 20201021_normalizePairCounts}

pairs.norm <- pairs %>% group_by(item1) %>% 
  arrange(desc(n), .by_group = TRUE) %>% 
  mutate(n_norm = (n / first(n)) %>% round(4), .after = n)

# TODO [] 2020-10-23 | Arrange by all, unique, shared for each group in item1 (factor bins?)
{  x <- pairs$item2
   x <- factor(x, levels = c(x[str_ends(x, "all")], x[str_ends(x, "unique")], "shared", 
                                          unique(pairs$item1)
                                   )
               )
  pairs.norm %>% ungroup %>% mutate(item2 = x) %>%  arrange(item1, item2)
}

## Add group metadata and region-specific colors
groups.metadata  <- df.markers %>% distinct(group, celltype, region, stage) %>%
                left_join(colors.area %>% select(category, color.region), by = c("region" = "category"))

colors.area <- colors.area %>% mutate(category = factor(category, 
                                                    levels = c("pfc", "motor", "somatosensory", 
                                                               "parietal", "msp", "temporal", "v1")))

groups.metadata$region %<>% factor(levels = levels(colors.area$category))



# Add region, stage and colors to `pairs`.
fuzzyjoin::regex_left_join()

pairs.norm %<>% left_join(groups.metadata,
                                        by = c("item1" = "group")
                                        ) %>%
                                regex_left_join(groups.metadata, 
                                                by = c("item2" = "group")) %>%
                                        # suffix = c(".source", ".target")
                               select(-group) %>% rename_with(.fn = ~case_when(str_detect(.x, ".x") ~ gsub(".x","_source", .x),
                                                                               str_detect(.x, ".y") ~ gsub(".y","_target", .x),
                                                                              TRUE ~ .x)
                               )

# function(x) { gsub(".x", "_source", x, fixed = TRUE)
#              gsub(".y", "_target", x, fixed = TRUE)
#                  
#                } )

# Check levels of categorical variables: stage, region
pairs.norm %>% map(~ if ( is.factor(.x) ) print(levels(.x)) )

write_rds(pairs.norm, path = "../../sankey/out/sankey_markers_pairs_exnLineage.rds")
# Saved on 20201023
# TODO figure out how to write as a simple text file? 
  # Column 'genes' is a list of character vectors.
                  
# Go to gvisSankey from here

datatable(pairs, class = 'table-bordered table-condensed')

```

### Plot colors for each area / region
```{r}
### TODO fix this table so it has all the info about region, region colors.
### colors.area 

colors.area <- colors.area %>% mutate(area = factor(area, 
                                                    levels = c("pfc", "motor", "somatosensory", 
                                                               "parietal", "msp", "temporal", "v1")))

x <- pairs %>% left_join(df.markers %>% 
                           distinct(group, celltype, region, stage, color, color.rg), 
                        by = c("item1" = "group")
                        ) 


# Prepare tables
x <- distinct(df.markers, area, region, color.region, color.area)
x$area <- factor(x$area, levels = levels(colors.area$area))
x$region <- factor(x$region, levels = levels(colors.area$area))
x$color.region <- factor(x$color.region, levels = x$color.region %>% unique)

x <- x %>% arrange(region)

ggplot(x) + 
  geom_point(aes(x = area, colour = region, y = 0), size = 10) +
  scale_color_manual(values = x$color.region %>% as.character() %>% unique)

```

# . . . . . . . . . . . . . . . . . . . . .

# gvisSankey
```{r gvisSankey}
# Start with df.links
# df.links: A table with the # of shared gene markers between pairs of cell groups.
# Source and target are in both name and number form

# Rows: 144
# Columns: 9
# $ source.name       <chr> "shared_rg_msp_early", "shared_rg_msp_early", "shared_rg_msp_early", "shared_rg_msp_early"…
# $ target.name       <chr> "to_neuron_msp_early", "to_neuron_msp_late", "to_neuron_msp_mid", "to_neuron_pfc_early", "…
# $ link.group.source <chr> "msp", "msp", "msp", "msp", "msp", "msp", "msp", "msp", "msp", "msp", "msp", "msp", "v1", …
# $ link.group.target <chr> "msp", "msp", "msp", "pfc", "pfc", "pfc", "temporal", "temporal", "temporal", "v1", "v1", …
# $ n                 <dbl> 113, 71, 116, 31, 27, 72, 13, 40, 47, 47, 16, 79, 151, 174, 264, 54, 62, 119, 74, 97, 121,…
# $ source            <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, …
# $ target            <dbl> 12, 20, 16, 14, 21, 19, 15, 23, 17, 13, 22, 18, 12, 20, 16, 14, 21, 19, 15, 23, 17, 13, 22…
# $ color.rg          <chr> "#FFA500", "#FFA500", "#FFA500", "#FFA500", "#FFA500", "#FFA500", "#FFA500", "#FFA500", "#…
# $ color             <chr> "orange", "orange", "orange", "#e72718", "#e72718", "#e72718", "#ff61e8", "#ff61e8", "#ff6…

gvis.df <- df.links %>% filter(str_detect(.$source.name, "pfc")) %>%
  
                                       arrange(source.name, target.name) %>% 
                                       select(source.name, target.name, n)


gvis.df <- df.links %>% filter(across(.cols = contains("name"),
                                        .fns = ~str_detect(., "late"))) %>%
                                      arrange(source.name, target.name) %>% 
                                       select(source.name, target.name, n)




```


```{r 20201022_make_gvis.dfs}

comparisons <- tribble(
                  ~celltype_1, ~celltype_2,

                        "rg",  "neuron",
                        "rg",  "ipc",
                        "ipc", "neuron"
                )

comparisons <- tribble(
                  ~celltype_1, ~celltype_2,

                        "rg",  "rg",
                        "ipc",  "ipc",
                        "neuron", "neuron"
                )

gvis.df_self <-  purrr::pmap(.l = comparisons, .f = ~ pairs.norm %>% select(-genes) %>%
                # mutate(n = as.integer(n)) %>%  [not sure if necessary]
                                        filter(item1 %>% str_detect(..1) & item2 %>% str_detect(..2)) %>% 
                                          filter(! item2 %>% str_detect("all|shared")) %>%
                                          arrange(celltype_source, stage_source, region_source, stage_target, region_target) %>% 
                                          mutate(item2 = str_remove(item2, "_unique"))) %>%
                                          tibble(item1_item2 = purrr::pmap_chr(.l = comparisons, ~ glue(..1, ..2, .sep = "_to_")), df = .)


```

```{r}
gvis.df_self$df
```


```{r}
sankey.same.celltype <- pmap(gvis.df_self, ~ make_gvisSankey(chart_id = ..1, gvis_df_x = ..2))


make_gvisSankey <- function(chart_id, gvis_df_x) {
  
  gvis_df_x %<>% mutate(item2 = paste0("to_", item2))
  # TODO fix color issue [NA is not a valid color string] 2020-10-23
  # remove to_ in item2 for making colors
  
  groups.order <- gvis_df_x %>% split(f = factor(.$item1, levels = unique(.$item1))) %>% 
                      map(~ c(unique(.$item1), .$item2)) %>% 
                          unlist(use.names = FALSE) %>% unique %>% 
                              tibble(group = . ) %>% 
                      mutate(group = str_remove(group, "to_"))
  
  groups.order <- left_join(groups.order, groups.metadata %>% select(-starts_with("color")))
  groups.order <- left_join(groups.order, colors.area, by = (c("region" = "category")))
                          
  colors <- paste("['", paste(groups.order$color.region, collapse = "\',\'"),"']", sep = "")
  
  
  
  
  plot <- gvisSankey( chartid = paste("sankey", chart_id, sep = "_"),
                data =  gvis_df_x %>% select(item1, item2, n), # Doesn't like having all those extra columns
                from = "item1", 
                 to = "item2", 
                 weight = "n",
                 options = list(width = 1000,
                                height = 800,
                                backgroundColor = "black",
                                tooltip = "{ 
                                      textStyle: { color: 'black' }, 
                                      showColorCode: 'True',
                                      isHTML: 'True'
                                      }",
                                
                                sankey = paste( "{ iterations: 0 ,
                                                         node: { colors : ", colors, ", 
                                                                width : 25,
                                                                nodePadding : 35,
                                                                interactivity: 'True'
                                                                              },
                                                      
                                                  label: { color: '#871b47' } , 
                                                  link: { colorMode: 'gradient' }
                          
                                                 }"
                                            )
                                        )
            )

print(plot, tag = "chart", file = glue("../../sankey/out/html/sankey_", chart_id, ".html")) 

return(list(groups.order, colors, gvis_df_x, plot) %>% tibble( chart_id = .))

}


options(op)
plot(x)
plot(x, tag = 'html')

htmltools::save_html(x$html$chart, file = "../sankey/out/test.html")
print(x, tag = "chart", file = "../sankey/out/test.html") 


op <- options(gvis.plot.tag = 'html')
op <- options(gvis.plot.tag = 'chart')
op <- options(gvis.plot.tag = NULL)

getOption("gvis.tags")

print(x, tag = 'chart')

op <- options(gvis.print.tag = 'html')
# https://stackoverflow.com/a/18236653
print(x, tag ="chart", file="../sankey/out/test.html") 




# Get groups in order in which they appear

groups.order <- gvis.df[[1]] %>% split(f = factor(.$item1, levels = unique(.$item1))) %>% 
                      map(~ c(unique(.$item1), .$item2)) %>% 
                          unlist(use.names = FALSE) %>% unique %>% 
                              tibble(group = .)

# . . . . . . . . . . . . . .
# Alternative ways I tried:
# https://github.com/tidyverse/dplyr/issues/4531

# gvis.df[[1]] %>% group_by(item1) %>% group_map(~ item1 %>% tibble())

# gvis.df[[1]] %>% nest(gg = -item1) %>% mutate_at("gg",
#                                                 purrr::map,
#                                                 function(x) x$item2)


# Pair with colors
colors.area <- read_tsv("../../colors_area.tsv")
groups.order <- left_join(groups.order, groups.metadata %>% select(-starts_with("color")))
groups.order <- left_join(groups.order, colors.area, by = (c("region" = "category")))
                          


colors = paste("['", paste(groups.order$color.region, collapse = "\',\'"),"']", sep = "")
"

paste( "{ iterations: 0 , node: { colors : ", colors, "},")



```

```{r}
sankey.same.celltype %>% str
```

               
```{javascript}

colors.area
colors: ['#ff2b1a', '#ff2b1a', '#ff2b1a', '#1c32ff', '#1c32ff', '#1c32ff' '#1c32ff','#1c32ff', '#1c32ff', '#1c32ff', '#1c32ff', '#1c32ff' '#1c32ff','#1c32ff', '#1c32ff']

colors: ['#FF2B1A', '#FF2B1A', '#FF2B1A', '#1C32FF', '#1C32FF', 
                                                      '#1C32FF' '#1C32FF','#1C32FF', '#1C32FF', '#1C32FF', 
                                                      '#1C32FF', '#1C32FF' '#1C32FF','#1C32FF', '#1C32FF']


colors: ['orange', 'orange', 'orange', '#e72718', '#e72718', 
         '#e72718', '#ff61e8', '#ff61e8', '#ff61e8', '#1b30f3', '#1b30f3', '#1b30f3', 
         'orange', 'orange', 'orange', '#e72718', '#e72718', '#e72718', '#ff61e8', '#ff61e8', '#ff61e8', '#1b30f3', '#1b30f3', '#1b30f3', 'orange', 'orange', 'orange', '#e72718', '#e72718',
          '#e72718', '#ff61e8', '#ff61e8', '#ff61e8', '#1b30f3', '#1b30f3', '#1b30f3', 
           'orange', 'orange', 'orange', '#e72718', '#e72718', '#e72718', '#ff61e8', '#ff61e8', '#ff61e8', '#1b30f3', '#1b30f3', '#1b30f3', 'orange', 'orange', 'orange',
           '#e72718', '#e72718', '#e72718', '#ff61e8', '#ff61e8', '#ff61e8', '#1b30f3', '#1b30f3', '#1b30f3', 'orange', 'orange', 'orange', '#e72718', '#e72718', '#e72718', '#ff61e8', '#ff61e8', '#ff61e8', '#1b30f3', '#1b30f3', '#1b30f3',
            'orange', 'orange', 'orange', '#e72718', '#e72718', '#e72718', '#ff61e8', '#ff61e8', '#ff61e8', '#1b30f3', '#1b30f3', '#1b30f3', 'orange', 'orange', 'orange', '#e72718', '#e72718', '#e72718', '#ff61e8', '#ff61e8', '#ff61e8', '#1b30f3',
                         '#1b30f3', '#1b30f3', 'orange', 'orange', 'orange', '#e72718', '#e72718', '#e72718', '#ff61e8', '#ff61e8', '#ff61e8', '#1b30f3', '#1b30f3', '#1b30f3', 'orange', 'orange', 'orange', '#e72718', '#e72718', '#e72718', '#ff61e8', '#ff61e8',
                         '#ff61e8', '#1b30f3', '#1b30f3', '#1b30f3', 'orange', 'orange', 'orange', '#e72718', '#e72718', '#e72718', '#ff61e8', '#ff61e8', '#ff61e8', '#1b30f3', '#1b30f3', '#1b30f3', 'orange', 'orange', 'orange', '#e72718', '#e72718', '#e72718',
                         '#ff61e8', '#ff61e8', '#ff61e8', '#1b30f3', '#1b30f3', '#1b30f3'] 

# Colors for PFC rg -> all neurons
colors: ['#FF2B1A',
'#e4840f',
'#e4840f',
'#e4840f',
'#e72718',
'#e72718',
'#e72718',
'#ff61e8',
'#ff61e8',
'#ff61e8',
'#1b30f3',
'#1b30f3',
'#1b30f3',
'#FF2B1A',
'#FF2B1A']
```

```{r}
colors %<>% tibble(area = names(colors), color = .)

# # A tibble: 4 x 2
#   area     color  
#   <chr>    <chr>  
# 1 pfc      #F94171
# 2 msp      #FFA750
# 3 temporal #3182BD
# 4 v1       #9E9AC8

colors.area$color.rg <- shades::brightness(colors.area$color, 1.5) %>% as.character()

df.links <- links %>% select(contains("name"), contains("link.group"), n, everything()) %>% 
  left_join(colors.area %>% select(category, color.rg), 
            by = c("link.group.source" = "category")) %>% 
                          left_join(colors.area %>% select(category, color), 
                                    by = c("link.group.target" = "category"))
                                                        
write_tsv(df.links, "../sankey/out/rg_to_neurons_df.tsv")
```


### networkD3
```{r}
sankey <- networkD3::renderSankeyNetwork(expr = nodePadding = 20,
                              # margin = list(top = 100, bottom = 100),
                              sinksRight = FALSE,
                              # height = 250,
                              # width = 500,
                    Links = links,
                    Nodes = nodes, 
                    Source = "source",
                    Target = "target", 
                    Value = "n", 
                    NodeID = "name",
                    units = "genes", # optional units name for popups
                    NodeGroup = "node.group",
                    LinkGroup = "link.group",
                    fontSize = 12, 
                    nodeWidth = 30,
                    iterations = 0,
                    colourScale = 'd3.scaleOrdinal().range(["#F94171", "#FFA750", "#9E9AC8", "#3182BD"])'
                   )
```

```{r}
stash <- rlang::env(markers = markers)

markers.plot.df <- markers.plot.x

as.tibble(markers.plot.x)
```

```{r}
markers.plot.x %>% plot_bar()

markers.plot %>% plot_bar()

inspectdf::inspect_cat(markers.plot) %>% show_plot(label_thresh = 0, label_size = 3, plot_type = 2)

markers.plot.x %>% skim

inspectdf::inspect_na(markers.plot) %>% show_plot

markers.plot[!complete.cases(markers.plot),] 
```


```{r}
inspectdf::inspect_cat(markers.plot.x %>% select(area, region, individual, stage)) %>% show_plot(label_thresh = 0, label_size = 3, plot_type = 2)
```

```{r}
inspectdf::inspect_num(markers.plot.x) %>% show_plot()
```

```{r}
pairs.raw
```

```{r}
pairs.raw %>% plot_bar()

plot_bar

dfSummary(pairs.raw)

write_tsv(pairs.raw, "../sankey/out/pairs_raw_rg_to_neurons.tsv")
```

```{r}
# [[Project: Sankey Plots]] #Roam
# https://roamresearch.com/#/app/carmen-se/page/knEV4Gm_H

x <- df.links %>% arrange(source.name, target.name) %>%  
  
  dplyr::filter(across(.cols = contains("name"),
                                        .fns = ~str_detect(., "PFC"))
)

x <- df.links %>% arrange(source.name, target.name) %>%  
  
  dplyr::filter(across(.cols = contains("name"),
                                        .fns = ~str_detect(., "PFC"))
)
  
sankey.plot <- with(x, ggSankeyGrad::ggSankeyGrad(
                           c1 = source.name, 
                           c2 = target.name, 
                           col1 = color.rg, 
                           col2 = color, 
                           values = n, label = TRUE, alpha = 0.75, padding = 50, color_steps = 50) 
) 



%>% ggplotly()
     


                           # label = TRUE, padding = 50, alpha = 0.75)

jsonedit(sankey.plot$)

x %>% jsonedit()

Sys.getenv(sankey.plot$layers[[1]])


df.links
```
```{r}
p_load(DataExplorer)
df.links %>% plot_str()
```

```{r}
df.links %>% plot_bar()
df.links %>% DataExplorer::plot_scatterplot(by = "source.name")
```

```{r}
df.links %>% DataExplorer::plot_density(geom_density_args = list(aes(colour = source.name)))

df.links %>% ggplot() + geom_point(aes(x =target.name, y = source.name, size = n, color = color.rg))
```

```{r}
df.links <- df.links %>% left_join(markers[ , c("")],  ...))

stages <- str_c(levels(markers.plot$stage), collapse = "|")
df.links <- df.links %>%  mutate(stage = .$source.name %>% str_extract(stages))


pairs %>% left_join(df.markers %>% select(group, celltype, region, stage), by = c("item1" = "group"))
pairs %>% ggplot() + geom_density(aes(x = n, colour = factor(stage))) + facet_wrap(~ color.rg)
```

Run quantification function on this.
```{r}
df.linksq
```

# / / / / / / / / / / / / / / / / / / / / / / / / / / / / / / / / / / 


## Make all Sankey plots for all stages.

## Prepare `markers` df for pairwise intersection calculation.
(Prepare for buildSankey)
```{r}
# Add column with cell type.

# markers.ct %<>% purrr::imap( ~ .x$all.indiv %>% mutate(cell.type = .y))

# markers %>% split(markers$individual)

      # Plot only early / mid / late stage links. (filter)

prepareMarkers <- function(markers.df = markers.df) {
                
                 # Filter by p-value
                 markers.df %>% dplyr::filter(p_val_adj <= 0.05) %>%
                                
                  # Optional: Combine levels 
                    # Combine ages and areas into meta-areas (regions) & meta-ages (stages)
                  mutate(individual = as_factor(individual),
                         
                         region = fct_collapse(area, 
                                                   msp = c("motor", "somatosensory", "parietal")
                                                   ),
                                  
                         stage = fct_collapse(individual,
                                                early = c("gw14", "gw16", "gw17"),
                                                mid = c("gw18_2", "gw18", "gw19_2", "gw19", "gw20_31and34", "gw20"),
                                                late = c("gw22both", "gw25")
                                               )) %>%
    
                      # Merge columns for easier manipulation
                      unite("area.individual", area, individual, sep = "_", remove = FALSE) %>%
                      unite("region.stage", region, stage, sep = "_", remove = FALSE) %>%
                      unite("group", celltype, region.stage, sep = "_", remove = FALSE) %>%

                      select(gene, group, celltype, region, stage, region.stage, area, individual, area.individual, everything() )
}


```


```{r prepareMarkers}
df.markers <- df.markers %>% prepareMarkers()
# This still has duplicate 

# df.markers has duplicate genes which may have appeared for a given group several times.
df.markers %>% select(-file, -color, -color.rg) %>% distinct(gene, celltype) %>% plot_bar

df.markers %>% select(-file, -color, -color.rg) %>% distinct(gene, region) %>% plot_bar

df.markers %>% select(-file, -color, -color.rg) %>% distinct(gene, individual) %>% plot_bar

df.markers %>% select(-file, -color, -color.rg) %>% distinct(gene, stage) %>% plot_bar

df.markers %>% select(-file, -color, -color.rg) %>% distinct(gene, area) %>% plot_bar

df.markers %>% select(-file, -color, -color.rg) %>% distinct(gene, region.stage) %>% plot_bar

df.markers %>% select(-file, -color, -color.rg) %>% distinct(gene, group) %>% plot_bar


# group : celltype:region:stage (egion:stage by cell type)

# getDistinct 
# [] Turn into a function
df.markers.distinct <- df.markers %>% 
                    # dplyr::filter(celltype == "neuron" | celltype == "rg") %>%
  
                      # Collapse genes that appear > 1x in a group (eg genes in neuron:early:msp)
                      distinct(gene, group, .keep_all = TRUE) %>% as.tibble()
                                             #  .keep.all: If TRUE, keep all variables in .data
```


```{r buildSankey_df.markers.distinct}

x <- buildSankey(markers = df.markers.distinct)

unique(na.exclude(markers$celltype)
      
       
markers.split <- markers.plot %>% split(.$celltype) 

sankey.plots.all <- lapply(markers.split[c("rg", "neuron", "ipc")], function(x) {
  
  # message(x)
  markers.plot.x <- x %>% filter(p_val_adj <= 0.05) %>%
                        distinct(gene, meta.area.age, .keep_all = TRUE)

stages <-  list(
  all = buildSankey(markers = markers.plot.x, stage = NULL, dir = "../out/all_areas"),
  early = buildSankey(markers = markers.plot.x, stage = "early", dir = "../out/all_areas"),
  mid = buildSankey(markers = markers.plot.x, stage = "mid", dir = "../out/all_areas"),
  late = buildSankey(markers = markers.plot.x, stage = "late", dir = "../out/all_areas")
  )
  return(stages)
}
  )

write_rds(sankey.plots.all, "../sankey/out/sankey_plots.rds")

df.markers.distinct %>% skim
```










# ..........................................

## Quantify Sankey plots

```{r}
listviewer::jsonedit(sankey.plots.all)


heatmaps.allareas <- lapply(names(sankey.plots.all), function(x) {
  
  pairs.df <- sankey.plots.all$x
  
  x  <- buildSharedGeneMatrix(cell.type = x, x = pairs.df, )
   
}



p.all <- patchwork::wrap_plots(purrr::map(heatmaps.2, 2), nrow = 3, guides = "collect")
pdf("../out/shared_genes_matrixes_identityDiag.pdf", width = 11 * 2.5, height = 8 * 2.5)
print(p.all)
dev.off()


heatmaps.allareas[[2]][[1]][[1]]$heatmap %>% plotly::ggplotly()

```

### buildSharedGeneMatrix

Builds the pair-wise shared gene matrix (number of genes in each link).

- Input: 
  - A `pairs.raw matrix` [in a sankey.plot object]
  
```{r}
markers.plot %>% plot_bar
```
  
```{r}
markers.plot.rg.neuron %>% plot_bar
```
```{r}
x <- gvis.df
x <- df.links
x <- markers.plot.rg.neuron
# Need the matrix that has all the intersections before adding to_ shared_ etc and with diagonal.
```

```{r buildSharedGeneMatrix}

buildSharedGeneMatrix <- function(x) {
                         # x = sankey.object$pairs.raw
  
     # pairs.matrix <- x$pairs.raw %>%
            x %>% pivot_wider(id_cols = 1, names_from = 2, values_from = n) %>% glimpse
              select(-item1) %>% as.matrix()

    diag <- diag(pairs.matrix)
    
    # unique.genes <- x[[stage]]$unique.genes$counts
    # unique.genes <- unique.genes[colnames(pairs.matrix), ]
    # diag(pairs.matrix) <- unique.genes
    
    pairs.matrix.norm <- apply(pairs.matrix, 2, function(x)  x/diag)

    # For ggplot-style plotting
    pairs.matrix.long <- pairs.matrix.norm %>% as.data.frame %>% 
                            set_rownames(names(.)) %>% rownames_to_column("item1") %>% 
                              pivot_longer( ! item1, names_to = "item2", values_to = "pct.shared")
    
    
    # Make matrix for heatmaply()
    df <- pairs.matrix.norm %>% as.data.frame %>%

             mutate(across(.cols = everything(), .fns = ~ round(.x, 2)),
                    stage = str_extract(colnames(.), "early|mid|late"), 
                     area = str_extract(colnames(.), colors.area$category %>% collapse(sep = "|"))
                    ) %>% 
                set_rownames(value = names(.[1:nrow(.)]))
                             
# df.meta <- df %>% left_join(colors) %>% select(stage, area, color)

return(tibble::lst(df, df.meta, pairs.matrix, pairs.matrix.long, pairs.matrix.norm))

}
```



```{r}
result <- sankey.plots.all %>% map_depth(.f = buildSharedGeneMatrix, .depth = 2)
```

```{r fxn-heatmaply}
# Actually make heatmap / dotplot
buildHeatmap <- function(x) {
  
              df <- x$df
              
              heatmap <- heatmaply(df[1:nrow(x$df)], 
                
              node_type = 'scatter', point_size_mat = df,
              scale = "none", 
              point_size_name = "% markers shared",
              Colv = "Rowv",  show_dendrogram = c(TRUE, FALSE), k_row = 3,
              row_side_colors = df[ , c("area", "stage")],
              row_side_palette = colors$color,
              scale_fill_gradient_fun = scale_colour_gradientn(
                                                               na.value = "white",
                                                               colors = bpy.colors(n = 100, alpha = 0.9),
                                                               values = c(0, 0.3, 1),
                                                                breaks = c(0, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9),
                                                                limits = c(0.01, 0.9),
                                                                labels = c(0, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9) %>% as.character,
                                                                name = "% of markers shared",
                                                                oob = scales::squish
                                                                ),
          heatmap_layers = list(scale_size_binned_area(# max_size = 10,
                                                # n.breaks = 100,
                                                breaks = c(0.0, 0.049, seq(0.05, 0.9, by = 0.1)), 
                                                limits = c(0.01, 0.9),
                                                oob = scales::censor
                            ))
                          )

# Resize widget so to a specific width/height and turn off filling entire browser.
# https://stackoverflow.com/questions/47921757/how-to-resize-html-widget-using-savewidget-in-htmlwidgets-r
heatmap$sizingPolicy$defaultWidth <- 800
heatmap$sizingPolicy$defaultHeight <- 450
heatmap$sizingPolicy$viewer$fill <- FALSE
heatmap$sizingPolicy$browser$fill <- FALSE

return(heatmap)
}


heatmaps <- result %>% map_depth(.f = buildHeatmap, .depth = 2)


htmlwidgets::saveWidget(widget = x, 
                        file = "../out/sankey_sharedGenes_heatmap.html", 
                        title = "sankey_sharedGenes_heatmap", selfcontained = TRUE)

combineWidgets(heatmaps$rg$all, heatmaps$neuron$all, heatmaps$ipc$all, nrow = 1)


# TODO use index as title of plot (eg. rg early)
# TODO separate parietal from motor, somato. or motor, somato, parietal
                        
```


# ________________________________________________________________________________________________

# Other stuff I tried
```{r}
split.df <- markers %>% group_by(area.individual) %>% dplyr::group_split()
  
intersections <- lapply(split.df, function(df.1 = x ) {
  
  lapply(split.df, function(df.2 = x) {
  
  genes <- intersect(df.1$gene, df.2$gene)
  n.intersect <-genes %>% length
  
  return(list(genes, n.intersect))
  }
  )
}
)

intersections %>% glimpse

#####

markers %>% pivot_wider(id_cols = gene, names_from = area.individual, values_from = area.individual) %>% group_by(.[,1]) %>% groups
  summarise(lapply(names(x)[1:2]), function(x) count(markers[,1], x))
  
#####
markers %<>% add_column(cell.type = "neuron")

# https://rdrr.io/cran/widyr/man/pairwise_count.html
pairs <- widyr::pairwise_count(markers, item = area.individual, feature = gene, 
                               sort = TRUE, diag = FALSE, upper = FALSE) %>% 
          arrange(item1, item2)

pairs %<>% select(item1, n, item2) %>% dplyr::filter(n >= 3)  

makeSankeymatic <- function(pairs.df, cell.type) {
  
  pairs.df %<>% mutate( n = pairs.df %>% str_glue_data("[{.$n}]"),
                    item2 = pairs.df %>% str_glue_data("to_{.$item2}")) %>%
    select(item1, n, item2)
  
    pairs.df %>% write_tsv(paste0("../out/", dir, "/pairwise_intersect_markers_", cell.type, ".tsv"), col_names = FALSE)
    
  return(pairs.df)
}

```


## Other stuff I had tried

```{r}

p <- makeHeatmap(df = pairs.matrix.long)


cell.type, 

results <- lapply(names(x), function(stage) {
  
                  
      
      return(list(matrix = pairs.matrix,  norm.matrix = pairs.matrix.norm, matrix.long = pairs.matrix.long, heatmap = p))
    


# Extract the 4th element of each list element.      
plots <- purrr::map(results, 4)

plots <- patchwork::wrap_plots(plots, nrow = 1, guides = 'auto')

return(list(results = results, plots = plots))

}

# p <- pairs.matrix.long %>% ggplot() + geom_jitter(aes(x = item1, y = pct.shared, colour = item2))


# TODO
# make diagonal pct of total that is unique
# ✔ change scale

```

```{r}


colors.bpy <- c("#000033FF", "#000099FF", "#0000FFFF", "#5000FFFF", "#9F0FF0FF", "#EF42BDFF",
                "#FF758AFF", "#FFA857FF", "#FFDB24FF" ,"#FFFF60FF")

colorRamp(colors = colors.bpy, bias = 10)(19)
scales::show_col(colors.bpy, ncol = 10, cex_label = 0.25)

bpy.colors(100) %>% scales::show_col(cex_label = 0.4)
colorRamp(colors = colors.bpy)(100)

```

```{r}
stage <- "early"
makeHeatmap(heatmaps.allareas[[2]][[1]][[1]]$matrix %>% as.data.frame())
heatmaply(heatmaps.allareas[[2]][[1]][[1]]$matrix %>% as.data.frame(), scale = "row")

p_load(heatmaps.allareas[[2]][[1]][[1]]$matrix.long %<>% mutate(across(.cols = starts_with("item"),
                                                                .fns = ~str_remove(., pattern = "_early|_mid|_late")
                                                               )
                                                        
```

```{r fxn: makeHeatmap}


makeHeatmap <- function(df = pairs.matrix.long) {
  
  if(!stage == "all") {
  
  df %<>% mutate(across(.cols = starts_with("item"),
                        .fns = ~str_remove(., pattern = "_early|_mid|_late")
                                                               )
                                                        )
  }
  
  heatmap <- ggplot(df) +
         
      geom_point(aes(x = item1, y = item2,
                     colour = round(pct.shared, 3),
                     size = round(pct.shared, 3)
                     ),
                     hjust = 0.5, vjust = 0.5
                 ) +
               # colour = "white") +
      scale_fill_gradientn(aesthetics = c("fill", "colour"),
                                          na.value = "white",
                           colors = bpy.colors(n = 100, alpha = 0.9), 
                           values = c(0, 0.3, 1),
                           limits = c(0, 0.99),
                           # labels = c(0, 0.1, 0.25, 0.5, 0.75, 0.8, 0.9, 1) %>% as.character,
                           name = "% of markers shared"
                           # oob = scales::squish
                           ) +
  scale_size_area(name = "% of markers shared") +
                  # labels = c(0, 0.1, 0.25, 0.5, 0.75, 0.8, 0.9, 1) %>% as.character) +
    
     ggtitle(paste(cell.type, stage)) +
     theme_minimal() +
     theme(line = element_blank(),
           axis.text.x = element_text(size = 10, angle = -45,
                                      hjust = 0, 
                                      vjust = 0.5
                                      # colour = colors.x, 
                                     # margin = margin(t = 0)
                                     ),
           axis.title = element_blank(),
           aspect.ratio = 1,
           axis.ticks = element_blank()
      )
  
      colorbar <- heatmaply::heatmapggplot_side_color_plot(
                    df,
                    palette = NULL,
                    scale_title = paste(type, "side colors"),
                    type = c("column", "row"),
                    text_angle = if (type == "column") 0 else 90,
                      is_colors = FALSE,
                    fontsize = 10,
  label_name = NULL
)
      
#geom_bar(mapping = aes(x = item1, y = 1, fill = item1), 
#                           stat = "identity", width = 1) +
#             theme_void() +
#             theme(panel.spacing.x = unit(0.5, "mm"),
#                   legend.position = "none")
#   
       plot <- plot_grid(colorbar, heatmap,
                         align = "v", ncol = 1, 
                        
                         axis = "l", rel_heights = c(0.5, 15)
                         )
       
       return(plot)
}