# Gene marker tables:
## df_markers_by_area_celltype_indiv.txt
    p-value goes from 0-1 (not filtered.)
    [2020-11-19] Combined and saved from files imported from:
    `aparna/homefiles/Fall2019_finalclustering_secondtrimester/_neocortex_*_byarea_clustermarkers.txt`
    Result of `FindMarkers` for each area vs everything else.
        - with cells of the same cell type
        - within each individual.
     See `sankey_plots.Rmd`

# Seurat object subsets:

```{r}
subsetSeurat(Neocortex, cells.keep = ncx.clusters.exn$cell.name %>% sample(50000), write.to = "../data/ncx.v3.exn.sub_50k.rds")
subsetSeurat(Neocortex, cells.keep = ncx.clusters.exn$cell.name %>% sample(100000), write.to = "../data/ncx.v3.exn.sub_100k.rds")
```

• Smaller dataset for working locally / not on massive cluster:
```{r}
# Custom function subsetSeurat (described below in fxns section)
subsetSeurat(Neocortex, "data/ncx_v3_40k.rds")

# On Mac
ncx.small <- read_rds("~/tmp/constellation_plots/c7104_ncx_v3_4k.rds")

# Reductions exported from neocortex_v3 [cfc4b2_neocortex_v3.RData] object:
# This is the only part that's really needed for the plots.
ncx.reductions <- read_rds("~/tmp/constellation_plots/11163_neocortex_v3_reductions.rds")
```

# Lean Seurat object subsets (rm @data, @counts):
```{r}
ncx.sub.lean <- ncx.sub %>% lapply(leanSeurat <- 
                                      function(x) {
                                        x@assays$RNA@counts <- matrix(c(0,0))
                                          x@assays$RNA@data <- matrix(c(0,0))
                                            return(x)
                                        }
                              )

write_rds(ncx.sub.lean$k100, "2ndTrimester/constellation_plots/data/ncx.v3.exn.sub_100k.rds")
write_rds(ncx.sub.lean$k50, "2ndTrimester/constellation_plots/data/ncx.v3.exn.sub_50k.rds")
```

# ______________________________________________________________________________
# Seurat
# ______________________________________________________________________________
## FindNeighbors
https://satijalab.org/seurat/v3.0/pbmc3k_tutorial.html
> We first construct a K-nearest neighbors graph based on the euclidean distance in PCA space.
  Refine the edge weights between any two cells based on the shared overlap in their local
  neighborhoods (Jaccard similarity). 
  This step is performed using the FindNeighbors function, and takes as input the previously defined
  dimensionality of the dataset (first 10 PCs).

```{r not-run}
FindNeighbors.default <- function(
  object,
  distance.matrix = FALSE,
  k.param = 20,
  compute.SNN = TRUE,
  prune.SNN = 1/15,
  nn.method = 'rann',
  annoy.metric = "euclidean",
  nn.eps = 0,
  verbose = TRUE,
  force.recalc = FALSE,
  ...
)
```

```{r}
ncx.small %<>% FindNeighbors(verbose = TRUE, do.plot = FALSE)
```

```{r}
ncx.small@graphs %>% lapply(corner)
```

# ______________________________________________________________________________
# Other functions
# ______________________________________________________________________________

## subsetSeurat
```{r}
# Custom function to downsample object.
subsetSeurat <- function(object = NULL, fraction = NULL, cells.keep = NULL, write.to = "path/to/filename.rds") {
  
                    if( is.numeric(fraction) ) { # Random subset of x% of all cells.
                       if(cells.keep == NULL) {
                        print("Subsetting to random subset of cells.")
                        cells.keep <- object@meta.data %>% rownames %>% sample(size = length(.)*fraction)
                       } else 
                         print("Either a fraction OR a vector of cells must be provided.")
                      
                    } else if( length(cells.keep) > 0) {
                        print("Using specified cells to create subset of Seurat object.")
                      
                    } else(print("Either a fraction OR a vector of cells must be provided."))
                    
                    obj.small <- object %>% subset(cells = cells.keep)
                    
                    if( ! write.to == "path/to/filename.rds" ) {
                      print("Writing file.")
                       write_rds(obj.small, write.to)
                    }
                     
                    return(obj.small)
              }
                        
```

## References
Install ggforce on cluster:
https://stackoverflow.com/a/59231055/4463919