---
title: "ChEA Analysis"
output: html_notebook
date: 2020-12-07
---

```{r}
dir_dropbox <- '~/Dropbox/ARK-lab/2nd Trimester Areas Paper/neocortex/markers/transcription_factors/ChEA'

p_load(VennDetail)
```

# Output from this notebook
```{r}
chea_table <- read_tsv(file.path(dir_dropbox, 'ChEA_TF_predictions_by_group.txt'))
```

# getChEA
```{r}
p_load(httr, js, jsonlite)

getChEA <- function(genes, query_name) {
  
		# genes = c("SMAD9","FOXO1","MYC","STAT1",'STAT3',"SMAD3")
									
									url = "https://maayanlab.cloud/chea3/api/enrich/"
									encode = "json"
									payload = list(query_name = query_name, gene_set = genes)
									
									#POST to ChEA3 server
									response = POST(url = url, body = payload, encode = encode)
									json = content(response, "text")
									
									#results as list of R dataframes
									results <- fromJSON(json)
									
	# Combine ChEA result
    # into a single table with a column for the `query_genes` for which a TF was predicted, 
    # and a column for the orginal source table.

			chea_table <- results %>%
                      map(~ rbindlist(.x, fill = TRUE, idcol = 'table')) %>% 
          
			
		# write_rds(file.path(dir_dropbox, 'ChEA_TF_predictions_by_group.rds'))
    # write_tsv(chea_table, path = file.path(dir_dropbox, 'ChEA_TF_predictions_by_group.txt'))
}
```

# Markers from each group
(cell type / area / stage)							
```{r}
markers <- read_rds('../../general/markers_MAST_wilcox_p0.1.rds')

markers %<>% imap(~ mutate(.x, test = .y)) %>% reduce(bind_rows)

markers_split_by_group <- markers %>% 
  unite(group, 'cell_type', 'stage', 'region', remove = FALSE) %>% split(f = .$group)

```

# ChEA-predicted TFs from each neuron area_celltype_stage group
Combine all groups into a single table.
```{r 2020-12-09}
chea_table <- markers_split_by_group %>% 
                imap(~ getChEA(genes = .x$gene, query_name = .y)) %>%
                  rbindlist() %>% 
                    rename(query_genes = 'Query Name') %>%
                      set_names(value = tolower(names(.))) %>%
                        separate(query_genes, 
                                 into = c('cell_type', 'stage', 'region'), 
                                 remove = FALSE) %>%
                           mutate(cell_type = factor(cell_type, levels = c('rg', 'ipc', 'neuron')),
                                  region = factor(region, levels = c('pfc', 'msp', 'temporal', 'v1')),
                                  stage = factor(stage, levels = c('early', 'mid', 'late')),
                                  score = as.numeric(score),
                                  ) %>% arrange(cell_type, region, stage)
```

# TF markers from each area_celltype_stage group
```{r}
tfs$MAST_wilcox_by_group %<>% ungroup %>%
  mutate(region = fct_collapse(area, 
                               msp = c('motor', 'somatosensory', 'parietal')))

# 2020-12-15
write_rds(tfs, '../out/TFs_MAST_wilcox_p0.1.rds')
```


# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 

# Find instersection
between group and upstream cell population markers

## Examples: PFC early and V1 early
```{r 2020-12-08}
x <- chea %>% dfilter(group %>% str_detect('early_v1')) %>% .$result

a <- tfs$MAST_wilcox_by_group %>% dfilter(group == 'rg_v1_early') %>% .$genes %>% .[[1]]

pfc_early <- map(x, function(chea_results) {
  
        b <- chea_results$`Integrated--meanRank`$TF
        intersection <- a[a %in% b]
        chea_results$`Integrated--meanRank` %>% dfilter(TF %in% intersection)
        
})

v1_early <- map(x, function(chea_results) {
  
        b <- chea_results$`Integrated--meanRank`$TF
        intersection <- a[a %in% b]
        chea_results$`Integrated--meanRank` %>% dfilter(TF %in% intersection)
        
})
```

## Pooled cell type-areas
```{r 2020-12-09}
regions <- c('pfc', 'msp', 'temporal', 'v1')

venn_diagrams <- regions <- c('pfc', 'msp', 'temporal', 'v1') %>% 
                    map(~ cheaVenn(region_use = .x, frac_keep = 0.2))

# Print diagrams to a single file.
plots <- map(venn_diagrams, wrap_plots) %>% wrap_plots(nrow = 4)

ggsave(plot = plots, filename = file.path(dir_dropbox,'venn_chea_rg_markers_top20pct_all_regions.pdf'), 
                                                                   width = 20, height = 30, device = 'pdf')
```


2020-12-15
# Controls: Mis-matched areas
```{r}
cheaControl <- function()
regions_compare <- expand_grid(region_rg_markers = regions, region_chea = regions)

venn_diagrams <- regions_compare %>% pmap( ~ cheaVenn(region_use_markers = .x, region_use_chea = .y, 
                                                     df_chea = chea_table, frac_keep = 0.2,
                                                     df_marker_tfs = tfs$MAST_wilcox_by_group,
                                                     table_use = 'Integrated--meanRank'))

venn_diagrams %<>% set_names(paste(regions_compare$region_rg_markers, 
                                   regions_compare$region_chea, sep = '_'))

venn_quant_df <- venn_diagrams %>% lapply('[[', 'venn_detail_df') %>% rbindlist(idcol = 'operation')

# Plot % of RG marker TFs that were predicted by ChEA
venn_quant_df %>% dfilter(type == 'input') %>% separate(operation, into = c('rg_marker_tfs', 'chea_tfs')) %>% 
  
  ggplot() + geom_point(aes(x = rg_marker_tfs, y = pct_shared, colour = chea_tfs)) + facet_wrap(~name)

```

# Get intersection between all SHARED sets.
```{r}
# Union of all genes in the 'shared' subsets.
venn_quant_df %>% dfilter(name == 'Shared') %>% .$value %>% unlist %>% unique
```



## Get top genes for each region 
```{r}

    # a) Find top n TFs -------------------------------------------------------------------
    chea_split <- chea_table %>% dfilter(table == 'Integrated--meanRank' & 
                                   query_genes %>% str_detect(paste0('neuron_.*_'))) %>% 
                                      split(.$query_genes)
    
    # Top 20% genes per region-stage category. (12 categories)
    chea_top_20pct <- chea_split %>% map(~ mutate(.x, score = as.numeric(score)) %>%
                                top_frac(n = 0.2, wt = -score)) %>%
                                  rbindlist
    
    # For how many REGIONS (4) was each gene a top 20% ChEA TF?
    # For how many REGION-STAGE (query_genes) (12) combinations was each gene a top 20% ChEA TF?

    chea_top_20pct %<>% group_by(tf) %>% 
      mutate(n_regions = n_distinct(region),
              n_queries = n_distinct(query_genes)) %>% 
      arrange(tf) %>% select(n_regions, n_queries, everything()) %>% ungroup
    

    
    tables <- list(counts = list(), genes_dfs = list())
    react <- function(x){reactable(x, fullWidth = FALSE, resizable = TRUE, pagination = FALSE)}
    
    # TFs per region-stage
    tables$counts$top20_by_region_stage <- chea_top_20pct %>% ungroup %>% count(query_genes) %>% 
                                              react
    
    tables$counts$top20_n_queries <- chea_top_20pct %>% count(n_queries) %>% react
    tables$counts$top20_n_regions <- chea_top_20pct %>% count(n_regions) %>% react

    # Filter TFs
    
     # Queries <= 6
      top20_filter_queries_6 <- chea_top_20pct %>% dfilter(n_queries <= 6)
      
      # Counts
      counts <- chea_top_20pct %>% group_by(query_genes) %>% summarise(n_distinct(tf))# 733 total genes
      counts_filter <-  top20_filter_queries_6 %>% group_by(query_genes) %>% summarise(n_distinct(tf))
    
      tbl <- left_join(counts, counts_filter, by = 'query_genes', suffix = c('_union', '_filter')) %>% react

    # Regions <= 2
      chea_top_20pct_filter <- chea_top_20pct %>% dfilter(n_regions <= 2)
    
      # Counts
      counts <- chea_top_20pct %>% group_by(region) %>% summarise(n_distinct(tf)) # 733 total genes
      counts_filter <- chea_top_20pct_filter %>% group_by(region) %>% summarise(n_distinct(tf))
      
      tbl <- left_join(counts, counts_filter, by = 'region', suffix = c('_union', '_filter')) %>%
        left_join(counts_filter_query <- top20_filter_queries_6 %>% group_by(region) %>% summarise(n_distinct(tf)), by = 'region') %>%
        react

```
    
2020-12-15    
# Run Venn comparison with filtered ChEA TFs

# TFs in <= 2 regions
```{r}
chea_top_20pct_filter <- chea_top_20pct %>% dfilter(n_regions <= 2)

control_filter_regions_2 <- cheaVennControl(df_chea = chea_top_20pct_filter, 
                                     frac_keep = 1,
                                     regions_compare = regions_compare, 
                                     title_filter = 'ChEA TFs in 1 or 2 neuron gene set queries')

control_filter_regions_2$venn_diagrams
```





2020-12-16
# TFs in <= 6 queries
```{r}
control_queries_6 <- cheaVennControl(df_chea = top20_filter_queries_6, 
                                     frac_keep = 1,
                                     regions_compare = regions_compare, 
                                     filename = 'filter_queries_6',
                                     title_filter = 'ChEA TF Filter: in 6 or fewer neuron gene set queries')

ggsave(plot = control_queries_6$p, filename = '../out/control_venn_top20ChEA_filter_queries_6.pdf', width = 70, height = 50, units = 'cm')
```

```{r}
control_top20 <- cheaVennControl(df_chea = chea_table, 
                                     frac_keep = 0.2,
                                     regions_compare = regions_compare, 
                                     title_filter = 'ChEA TFs: top 20% in each gene set query, pooled')


```


```{r}
control_top20_b <- cheaVennControl(df_chea = chea_top_20pct, 
                                     frac_keep = 1,
                                     regions_compare = regions_compare, 
                                     title_filter = 'ChEA TFs: top 20% in each gene set query, pooled')



# IDENTICAL to getting top 20 inside the function :)

```

# . . . . . . . . . . . . . . .
# Functions
```{r cheaVennControl}
cheaVennControl <- function(df_chea,
                            frac_keep = 1,
                            df_marker_tfs = tfs$MAST_wilcox_by_group,
                            regions_compare,
                            title_filter,
                            filename) {
  
      venn_diagrams <- regions_compare %>% pmap( ~ cheaVenn(region_use_markers = .x, 
                                                            region_use_chea = .y, 
                                                            df_chea = df_chea,
                                                            frac_keep = frac_keep,
                                                            df_marker_tfs = df_marker_tfs,
                                                            table_use = 'Integrated--meanRank'))
      
      venn_diagrams %<>% set_names(paste(regions_compare$region_rg_markers, 
                                         regions_compare$region_chea, sep = '_'))
      
      venn_quant_df <- venn_diagrams %>% lapply('[[', 'venn_detail_df') %>% 
        rbindlist(idcol = 'operation') %>% 
        separate(operation, into = c('input_markers', 'input_chea'), remove = FALSE) %>%
        rename(set = 'name')
      
      venn_diagrams <- map(venn_diagrams, ~ .x$plots$p_venn) %>%
                          wrap_plots(plots, nrow = 4)
      

      # Plot % of RG marker TFs that were predicted by ChEA
      p_df <- venn_quant_df %>% dfilter(type == 'input')
        
      p1 <- ggplot(p_df %>% dfilter(set == 'tfs_markers'), 
                   aes(x = input_chea, y = pct_shared, colour = input_markers, group = input_markers)) + 
        geom_point(size = 5, alpha = 0.5) +
        geom_line()  +
        theme_minimal() + ggtitle('% of radial glia markers in ChEA TF set')
      
     p2 <- ggplot(p_df %>% dfilter(set == 'tfs_chea'),
                  aes(x = input_markers, y = pct_shared, colour = input_chea, group = input_chea)) + 
        geom_point(size = 5, alpha = 0.5, ) +
        geom_line() +
        theme_minimal() + ggtitle('% of ChEA TFs in RG marker set')
      
     p_quant <- wrap_plots(plot_spacer(), p1, plot_spacer(), p2,  plot_spacer(), nrow = 5, heights = c(0.25, 1, 0.25, 1, 0.25))
     
     regions <- c('pfc', 'msp', 'temporal', 'v1') %>% as.list()
     
     text_regions_cols <- paste('ChEA TFs: ', regions) %>% map( grid::textGrob) %>% wrap_plots(nrow = 1)
     text_regions_rows <- paste('rg markers: ', regions) %>% map( grid::textGrob) %>% wrap_plots(nrow = 4)
     text_regions_rows <- plot_spacer() + text_regions_rows +
       plot_layout(nrow = 2, heights = c(0.1, 4))
     
       
     patchwork <- wrap_plots(text_regions_cols, venn_diagrams, nrow = 2, heights = c(0.1, 4))
     patchwork <- wrap_plots(text_regions_rows, patchwork, ncol = 2, widths = c(0.5, 4))
     
     p <- wrap_plots(patchwork, p_quant, nrow = 1) + 
       plot_layout(widths = c(5, 2)) + 
       plot_annotation(title = title_filter, theme = theme(plot.title = element_text(size = 12)))
     
    plot(p)
    ggsave(plot = p, filename = paste0('../out/control_venn_top20ChEA_', filename, '.pdf'), width = 75, height = 50, units = 'cm')
    return(lst(venn_diagrams, venn_quant_df, p))
}
```

```{r}


     
```


cat_n <- function(x) {paste0('\n', x, '\n')}



Add gene names to the venn diagrams
```{r}
p_load(RAM)


test %>% map(~ ggsave(plot = .x, filename = 'test.pdf', device = 'pdf'))

getwd()

p[[7]]$x %<>% .[1]
p[[6]]$y

p[[5]]$label

grid::grid.newpage()
grid::grid.draw(p)
      
region_use = 'pfc'
```


pdf('../out/venn/_test.pdf', height = height_p, width = width_p)
p <- plot_grid(p, labels = )
print(p)
dev.off()

# 1,655 TFs in whole TF db (humanTFdb).
# 1,632 TFs in ChEA



score_distrib <- df %>% as_tibble %>% ggplot(aes(x = as.numeric(score))) +
  geom_density(aes(fill = query_genes, colour = query_genes), alpha = 0.3) +
   geom_density(alpha = 0.2, colour = 'grey20', linetype = 'dashed', show.legend = FALSE) +
  theme_minimal()

ggsave(plot = score_distrib, file.path(dir_dropbox, 'tf_score_distribution_pfc_early.pdf'), width = 7, height = 5)  

plot <- chea_table %>% dfilter(table == 'Integrated--meanRank' & query_genes %>% str_detect('neuron_.*_pfc')) %>% 
  dfilter(tf %in% rg_markers_pool) %>% 
  scoreDistribution()


rg_markers_pool <- df_marker_tfs %>% unlist(x = .$genes, use.names = FALSE) %>% sort %>% unique
```

			