---
title: "TF Expression Dot Plots"
date: 2020-11-17
output: html_notebook
---

# Marker TF Dotplots: expression across cells from other areas . . . . . . . . . . . . .
```{r}
load(file = "./2020-11-15_transcription_factor_composition_workspace.RData")
sms("Loaded image")
```

# Add GW16 + GW19_2 to ncx.exn
```{r}
data.dir <- "~/cse-phd/second-trimester/neocortex/ncx_data/"

ncx.exn <- read_rds(paste0(data.dir, "exn_lineage/neocortex_exn_lineage_271k.rds"))
# 271K cells
load(paste0(data.dir, "gw16neo.RData"))
load(paste0(data.dir, "gw19_2.RData"))
gw16_cluster_ids <- read_tsv(paste0(data.dir, "gw16neo_clusteridentity.txt"))
gw19_cluster_ids <- read_tsv(paste0(data.dir, 'gw19_2_clusteridentity_wholebrain_annotated.txt'))

sms("Loaded objects")
```

# Clean up cluster annotation tables and add cluster info to metadata.

## Functions . . . . . . . . . . . . . . . 
```{r}
cleanClusterIDs <- function(cluster_ids) {
  
 cluster_ids <- cluster_ids %>% 
                  mutate(cell_type = cell_type %>% 
                                      str_replace("Radial Glia", "rg") %>% 
                                      str_replace("Excitatory Neuron", "neuron") %>%
                                       tolower)
  return(cluster_ids)            
}
```

```{r}
cleanMetadata <- function(meta.data, cluster_ids) {
    
                     meta.data <- meta.data %>%
                                  add_column(cell_name = rownames(.), .before = 1) %>%
                                    set_names(tolower(names(.))) %>%
                                      mutate(across(.cols = c(area, structure, individual), .f = tolower)) %>%
                                        left_join(cluster_ids) %>%
                                          set_rownames(.$cell_name)
                     
                     return(meta.data)
                     
                    }
```

```{r}
# gw19_cluster_ids %>% df_barPlot(variable = cell_type)
# gw16_cluster_ids %>% df_barPlot(variable = cell_type)

# gw16
gw16_cluster_ids %<>% cleanClusterIDs
gw19_cluster_ids %<>% cleanClusterIDs

gw16neo@meta.data %<>% cleanMetadata(cluster_ids = gw16_cluster_ids)
gw19_2@meta.data %<>% cleanMetadata(cluster_ids = gw19_cluster_ids)
```

## Merge samples with ncx.exn
```{r}
# [Satija Lab: Merge](https://satijalab.org/seurat/v3.2/merge_vignette.html)

# 1. Update gw16 and gw19_2 to v3
# 2. Subset only neocortex cells from gw19_2. (gw16 is only neocortex; Aparna)
# 3. Update ncx.exn
gw19_2_ncx <- UpdateSeuratObject(gw19_2) %>% subset(subset = `structure` == "neocortex")
gw16_ncx <- UpdateSeuratObject(gw16neo)
# Add counts back to ncx.exn!!!
ncx_counts <- read_rds(paste0(data.dir, "exn_lineage/ncx.v3.exn_counts.rds"))
ncx.exn@assays$RNA@counts <- ncx_counts
ncx.exn <- UpdateSeuratObject(ncx.exn) 

# Temporary .rdata file in case the merge crashes again:
save(gw16_ncx, gw19_2_ncx, ncx.exn, 
     file = paste0(data.dir, "ncx.exn_gw16_gw19_2_v3.RData"))



ncx_full <- merge(x = ncx.exn, y= c(gw16_ncx, gw19_2_ncx),
                 merge.data = TRUE,
                 project = "ncx_exn_full"
                 # add.cell.ids = c("gw19_2", "gw16") This appends to the actual cell_id (rowname)
              )
# Warning message:
#In CheckDuplicateCellNames(object.list = objects) :
#  Some cell names are duplicated across objects provided. Renaming to enforce unique cell names.

> ncx_full
# An object of class Seurat 
# 30250 features across 368806 samples within 1 assay 
# Active assay: RNA (30250 features, 0 variable features)
```
# Summarize metadata variables

```{r}
df_barPlot <- function(df, variable) { 
  
  variable <- enquo(variable)
  p <- df %>% ggplot + geom_bar(aes(y = !!variable))
  print(p)
}

var_plots <- quos(cell_type, individual, structure) %>% 
              map(~ df_barPlot(ncx_full@meta.data , variable = !!.x))

# [x] Put these all together in a single patchwork [2020-11-16]
wrap_plots(var_plots, nrow=2) + facet_wrap(~individual)

df_barPlot(ncx_full@meta.data , variable = cell_type) + facet_wrap(~individual)
```

Add _2 to gw19 cells $individual
```{r}
ncx_full@meta.data %<>% mutate(individual = case_when(cell_name %>% str_detect("gw19_2") ~ "gw19_2",
                                                                                    TRUE ~ individual))

gw19_2@meta.data %<>% mutate(individual = case_when(cell_name %>% str_detect("gw19_2") ~ "gw19_2",
                                                                                  TRUE ~ individual))

gw19_2@meta.data %>% filter(cell_name %>% str_detect("gw19_2")) %>% count(individual)
ncx_full@meta.data %>% filter(cell_name %>% str_detect('GW16')) %>% count(individual)
✅

ncx_full@meta.data %>% set_rownames(ncx_full@meta.data$cell_name)
# Error in `.rowNamesDF<-`(x, value = value) : duplicate 'row.names' are not allowed


gw19_2_ncx
# 43665 cells
ncx_full@meta.data %>% filter(cell_name %>% str_detect('gw19_2'))
# 52,293
ncx.exn@meta.data %>% filter(cell.name %>% str_detect('gw19_2'))
# It was already there!!! 😭
```

Fix metadata column names before merging objects.
```{r}
seurat_objects <- list(ncx.exn, gw16_ncx, gw19_2_ncx) %>% 
                      map(.f = function(x = .x) {
                        x@meta.data <- set_names(x@meta.data, 
                                                 str_replace_all(names(x@meta.data), 
                                                                "[.]", "_"))
                        return(x)
                        })

seurat_objects %>% map(~.x@meta.data %>% rownames %>% head)

seurat_objects[[1]]@meta.data %<>% set_rownames(.$cell_name)

metadata_common <- intersect(seurat_objects[[3]]@meta.data %>% names, 
                             seurat_objects[[1]]@meta.data %>% names)

seurat_objects[[1]]@meta.data %<>% .[ , metadata_common]

ncx_full <- merge(x = seurat_objects[[1]], y = seurat_objects[[2]],
                 merge.data = TRUE,
                 project = "ncx_exn_full"
                 # add.cell.ids = c("gw19_2", "gw16") This appends to the actual cell_id (rowname)
              )

ncx_full@meta.data %>% split(.$individual) %>% map(~ rownames(.x) %>% print)

ncx_full@meta.data %<>% mutate(individual = case_when(cell_name %>% str_detect("gw19_2") ~ "gw19_2",
                                                                                    TRUE ~ individual))


# FIXED:
  # ERROR Some cell names are duplicated across objects provided. Renaming to enforce unique cell names.
  # Error in names(new.idents) <- rownames(x = combined.meta.data) : 
  # 'names' attribute [596535] must be the same length as the vector [325141]
  # [x] What are the names of gw19_2 cells like? They include 'gw19_'.

# [seurat v3.0 merge two objects with different number of features | satijalab/seurat]
# (https://github.com/satijalab/seurat/issues/1676)
```

## Clean up metadata for ncx_full merged object and expand with the variables we need for plots.

## Expand metadata
```{r}
var_plots <- quos(cell_type, individual, structure) %>% 
              map(~ df_barPlot(ncx_full@meta.data , variable = !!.x))

# [x] Put these all together in a single patchwork [2020-11-16]
wrap_plots(var_plots, nrow=2) + facet_wrap(~individual)
# [x] [2020-11-16] NEXT: Fix metadata for merged object so that I can run the DotPlot by area / region / stage.
  # Go to DotPlot from here.
```


# Clean up and expand metadata
```{r}
# [x] 
  # [x] Change 'individual' to gw19_2 for cells with gw19_2 (currently gw19)
  # [x] cell_type tolower
  # [x] factor areas, regions, individuals, stages.
    # [x] Bin 'area' into 'region'
    # [x] Bin 'age' into 'stage'
  # [x] Make Neocortex -> neocortex
  # Subset only rg, ipc, neuron #nice-to-have

# From `sankey_plots.Rmd`

expandMetadata <- function(meta.data) {
  
            meta.data %>% set_names(names(.) %>% tolower) %>%
              mutate(cell_type = tolower(cell_type),
                     structure = tolower(structure),
                     individual = as.factor(tolower(individual)), 
                     # use base as.factor so it orders them numerically
                     stage = fct_collapse(individual,
                                          early = c("gw14", "gw16", "gw17"),
                                           mid = c("gw18_2", "gw18", "gw19_2", 
                                                   "gw19", "gw20_31", "gw20_34", "gw20"),
                                           late = c("gw22", "gw22t", "gw25")),
                     area = factor(tolower(area), 
                                   levels = c('pfc', 'motor', 'somatosensory', 
                                              'parietal', 'temporal', 'v1')),
                     region = fct_collapse(area, 
                                           msp = c("motor", "somatosensory", "parietal"))
                     ) %>%
                     unite("stage_region", stage, region, sep = "_", remove = FALSE) %>%
                     unite("stage_area", stage, area, sep = "_", remove = FALSE) %>%
                     unite("celltype_stage", cell_type, stage, sep = "_", remove = FALSE) %>%
                     unite("celltype_region", cell_type, region, sep = "_", remove = FALSE)

}

metadata <- ncx_full@meta.data %>% expandMetadata
metadata %>% sample_frac(0.1) # ✅ 
ncx_full@meta.data <- metadata


writeLines("[2020-11-17] Saved new version of ncx_exn with gw16 and updated metadata. 
             > write_rds(ncx_full, paste0(data.dir, 'exn_lineage/neocortex_exn_seuratobj.rds'))
             TODO Still has non-excitatory cells from gw16. Remove these.", 
             con = paste0(data.dir, 'exn_lineage/neocortex_exn_seuratobj.info'))
```

# Make dot plots . . . . . . . . . . . . . . . . .
# Set ident by 'celltype'
```{r}
p_load(viridis)
Idents(ncx_full) <- 'cell_type'
Idents(ncx_full) %>% table
p_load(sp)

  ncx_full@meta.data %<>% arrange(stage, region) %>%
    mutate(stage_region = as_factor(stage_region)) %>%
    arrange(stage, area) %>%
    mutate(stage_area = as_factor(stage_area))  
  
dotplots_rg_pfc_byStage <- imap(genes_list, 
                                ~ DotPlot(object = ncx_full,
                                          idents = 'rg',
                                          features = .x,
                                          group.by = 'stage',
                                          split.by = 'area',
                                          scale = TRUE,
                                          cols =  "RdYlBu",
                                          dot.min = 0.01) + 
                                  ggtitle(.y %>% str_extract("early|mid|late")) +
                                 
                                 scale_colour_gradientn(colors = bpy.colors(n = 100, alpha = 0.9)
                                                # TODO Maybe keep high yellow, low blue gradient.
                                                  # values = c(0, 0.3, 1),
                                                  # breaks = c(0, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9),
                                                  # limits = c(0.01, 0.9),
                                                  # na.value = "white"
                                 ) +
                                
                                scale_y_discrete(labels = function(x) str_extract(x, areas.rgx)) +
                                  
                                 theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
                                        axis.title = element_blank(),
                                        legend.position = "bottom") 
)

areas.rgx <- ncx_full@meta.data$area %>% unique %>% paste(collapse = "|")

str_split(ncx_full@meta.data$stage_area, n = 1, pattern = "_", simplify = TRUE)

# https://stackoverflow.com/questions/13649473/add-a-common-legend-for-combined-ggplots
# https://stackoverflow.com/questions/14604435/turning-off-some-legends-in-a-ggplot
patchwork <- wrap_plots(dotplots_rg_pfc_byStage, ncol = 3) + plot_layout(widths = c(6, 39, 7)) 

patchwork %<>% modify_at(.at = c(2,3), 
                          .f = ~ .x + theme(axis.text.y = element_blank(),
                                           axis.ticks.y = element_blank(),
                                           axis.title.y = element_blank(),
                                           ) +
                                     guides(size = FALSE)
                )

ggsave(plot = patchwork, filename = "../out/plots/dotplot_rg_pfc.pdf", width = 12.5, height = 6, units = 'in')

# Cannot install ggarrange
p_load(ggpubr)
ggarrange(dotplots_rg_pfc_byStage)

genes_list %>% map(length)


ncx.exn@meta.data$area
# TODO @issue Split.by does not respect factor order

ncx_full@meta.data %>% filter(cell_type == 'rg') %>%
  df_barPlot(cell_type) + facet_grid(rows = vars(stage), cols = vars(region))
```
