---
title: "Heatmaps: DE genes: cell types across brain regions (whole brain)"
output: html_notebook
---

# Data
```{r}
ncx.exn <- read_rds("../data/exn_lineage/neocortex_exn_lineage_271k_noCounts.rds")
```


# Brain region markers: by cell type
For each cell type, get the top marker genes for each area.

Read-in cell-type/areal marker genes

## Individuals separately
- Compare genes across individuals?
How consistent is the areal signature across age for a given cell type?
```{r}
dir <- "~/Dropbox/2nd Trimester Areas Paper/Area Specific Genes/"
files <- paste0(dir, list.files(dir)[list.files(dir) %>% str_detect("Neuron")])

# Use read.delim bc 1st column is rownames (tidyr functions don't work)
markers.ct <- list()

markers.ct$neuron <- lapply(files, function(x){
  
                 read.delim(x) %>% 
                  mutate(gene.ratio = pct.1 / pct.2,
                         gene.score = avg_logFC * gene.ratio
                        ) 
                  }) %>%
  
 set_names(value = files %>% tolower %>% 
          str_match(pattern = ".*/(.*)_neocortex.*") %>% .[,2]) %>%
  
           purrr::imap( ~mutate(.x, individual = .y))

markers.ct$neuron$all.indiv <- markers.ct$neuron %>% 
                                rbindlist()
```


## Exploration: gene scores, p_values, etc. 
```{r}
markers.ct$neuron$all.indiv %>% filter(cluster == "PFC") %>% group_by(individual) %>% 
  summarise(min(gene.ratio), max(gene.ratio), min(gene.score), max(gene.score), min(avg_logFC ), max(avg_logFC))
```

```{r}
markers.ct$neuron$all.indiv %>% filter(cluster == "PFC") %>% add_count(individual) %>% 
  ggplot() +
  geom_jitter(aes(individual, gene.ratio, colour = individual)) +
  scale_y_continuous(limits = c(0, 25)) +
  geom_label(aes(individual, 0, label = n))
```

```{r}
markers.ct$neuron$all.indiv %>% dplyr::filter(cluster == "PFC") %>% add_count(individual) %>%

  ggplot() +
  geom_jitter(aes(individual, avg_logFC, colour = individual)) +
  scale_y_continuous(limits = c(0, 2.75)) +
  geom_label(aes(individual, 0, label = n))


```

```{r}
genes$ %>% arrange(p_val_adj) %>%

  ggplot() +
  geom_jitter(aes(individual, gene.score, colour = p_val_adj, alpha = p_val_adj), size = 1) +
  scale_y_continuous(limits = c(0, 25)) +
  geom_label(aes(individual, 20, label = n)) +
  scale_color_gradient(low = "blue", high = "red", limits= c(0, 0.05)) +
  theme_minimal()

# Max is 50
```

```{r}
markers.ct$neuron$all.indiv %>% dplyr::filter(cluster == "PFC") %>% n_distinct(.$gene)
# 1,108

markers.ct$neuron$all.indiv %>% dplyr::filter(cluster == "PFC" & p_val_adj < 0.05) %>% n_distinct(.$gene)
# 981

markers.ct$neuron$all.indiv %>% dplyr::filter(cluster == "PFC" & p_val_adj <= 0.05) %>% add_count(individual) %>%

  ggplot(aes(x = individual, y = p_val_adj, colour = individual)) +
  geom_jitter(alpha = 0.5, size = 0.5) +
  geom_label(aes(individual, 0.025, label = n)) +
  theme_minimal()
  # scale_y_continuous(limits = c(0, 0.05))
  
```

## Genes per individual / per area
```{r}

colors.area$area %<>% factor(. , levels = .)
markers.ct$neuron %>% lapply(function(x) dplyr::count(x, cluster, individual))

markers.ct$neuron$all.indiv %>% dplyr::filter(p_val_adj <=0.05) %>% 
  dplyr::count(cluster, individual) %>%
  mutate(cluster = factor(cluster, levels = colors.area$area)) %>%
  
  ggplot() +
  geom_label(aes(cluster, individual, size = n, label = n, fill = cluster),
             color = "white", label.r = unit(0.3, "lines")
             ) +
  
  scale_x_discrete(breaks = colors.area$area) +
  scale_fill_manual(values = colors.area$color) +
  
  theme_minimal() + 

  guides(color = FALSE,
         size = FALSE,
         fill = FALSE) +
    
  theme(axis.text.x = element_text(angle = 35),
        text = element_text(size = 12),
        axis.title = element_blank()
        ) +
    
    ggtitle(label = "DE genes: Neurons")
  
  
  pivot_wider(id_cols = cluster, names_from = individual, values_from = n)
```

## Bin all genes into areas
```{r}

 genes <- markers.ct$neuron$all.indiv %>%  
          select(-X) %>% 
            dplyr::filter(p_val_adj <= 0.05) %>%
              group_by(cluster) %>% 
                group_modify(~ .x %>% group_by(gene) %>% 
                                       add_count(name = "n", sort = TRUE) %>% ungroup %>%
                                         select(gene, n, individual, everything())
                                       )  %>%
    group_split()

  # purrr::map(~ dplyr::count(.x , ..1 = cluster, ..2 = individual))

```


# Run clustering
otter.matrix <- as.matrix(otter.scaled[, -c(1:3)])
rownames(otter.matrix) <- otter.scaled$accession
otter.dendro <- as.dendrogram(hclust(d = dist(x = otter.matrix)))


## Make matrix for hierarchical clustering / dendrogram
```{r}
g <- genes[[6]]

g %<>% select(gene, individual, gene.score) %>% 
      pivot_wider(values_fill = 0,  
                  values_from = gene.score,  names_from = individual, id_cols = gene)
  
dd <- g %>% select(-gene) %>% as.matrix %>% set_rownames(g$gene) %>%
dist %>% hclust %>% dendro_data_k(9)
```

## Dendrogram
```{r}
cols <-  colors.use$cluster_color
  p <- plot_ggdendro(dd,
                   direction   = "tb", label.size = 1, scale.color = cols, branch.size = 0.5)
  
p

ggsave(p , filename = "../violinPlots/neocortex_DEgenes_by_celltype_area/dendrotest.pdf", height = 5, width = 20)

ggdendrogram(rotate = TRUE, size = 2) %>% ggsave(filename = "../violinPlots/neocortex_DEgenes_by_celltype_area/dendrotest.pdf", height = 11, width = 5)
```

# Heatmap: Gene scores
```{r}

g <-  g %>% pivot_longer( cols = 2:9, values_to = "gene.score",  names_to = "individual") %>% 
          dplyr::na_if(0) %>%
          mutate(gene.score = round(gene.score, 2)
                  # gene = factor(gene, levels = dd$labels$label)
          ) %>% left_join(genes$PFC %>% select(-X, -individual, -gene.score), by = "gene") 

g$gene <- factor(g$gene, levels = dd$labels$label)

```

```{r make-heatmaps}
genes$PFC.pval <- genes$PFC %>% dplyr::filter(p_val_adj <= 0.05)


heatmaps <- list()

heatmaps$range <- lapply(2:7, function(x) {
  
  print(x)
  df <- genes$PFC.pval %>% dplyr::filter(n >= x)
  
  p <- heatmapGenes(df = df, 
             color.by = gene.score, max.value = 5, 
             title = paste("Neurons: PFC: n >=", x, ":", n_distinct(df$gene), "genes")
  )
  return(p)
  
}
)

heatmaps$bin <- lapply(1:7, function(x) {
  
  df <- genes$PFC.pval %>% dplyr::filter(n == x)
  
  p <- heatmapGenes(df = df, 
             color.by = gene.score, max.value = 5, 
             title = paste("Neurons: PFC: n =", x, ":", n_distinct(df$gene), "genes")
  )
  return(p)
  
}
)



pdf(file = "../violinPlots/neocortex_DEgenes_by_celltype_area/heatmap_markers_neuron_PFC.pdf", onefile = TRUE, width = 30, height = 20)

patchwork::wrap_plots(heatmaps$range, ncol = 2, nrow = 4, guides = "collect", byrow = FALSE)
patchwork::wrap_plots(heatmaps$bin, ncol = 2, nrow = 4, guides = "collect", byrow = FALSE)
dev.off()

df <- g %>% dplyr::filter(n == 1)

heatmaps$single <-   heatmapGenes(df = df, 
             color.by = gene.score, max.value = 5, 
             title = paste("Neurons: PFC: n = 1", ":", n_distinct(df$gene), "genes")
)

# patchwork::wrap_plots(heatmaps, ncol = 3, guides = "collect", legend.position = "bottom", nrow = 3)
```


## fxn: HeatmapGenes  
```{r fxn-heatmapGenes}
heatmapGenes <- function(markers.df, n.genes, color.by, max.value, title) {

color.by <- enquo(color.by)

# Select genes that reoccur n times
g <- markers.df %>% dplyr::filter(n == n.genes) %>%
        mutate(gene.score = round(gene.score, 5))

dplyr::count(g, individual, wt = 1) %>% print

# Cluster genes
g %<>% select(gene, individual, gene.score) %>% 
          pivot_wider(values_fill = -3,  
                      values_from = gene.score,  
                      names_from = individual, 
                      id_cols = gene)


# Hierarchical clustering
# Get dist matrix, run hcust, extract ordered labels
dd <- g %>% ungroup %>% 
    dplyr::select(-gene) %>% as.matrix %>% set_rownames(g$gene) %>%
       dist %>% hclust %>% dendro_data

# Revert to dataframe for plotting
g <- g %>% pivot_longer(cols = 2:ncol(g), 
                          values_to = "gene.score",  
                          names_to = "individual") %>% 
          dplyr::na_if(-3) %>% 
          dplyr::filter(! is.na(gene.score))  %>%
            left_join(markers.df,
                      by = c("gene", "individual"))  %>%
mutate(gene = factor(gene, levels = dd$labels$label)) %>%
arrange(gene)

#levels <- g$gene %>% as.character %>% unique

#g %<>% mutate(gene = factor(gene, 
#                         levels = levels))

return(g)
}
```

```{r}
genes %>% names
```


```{r}
dfs <- lapply(1:max(genes$PFC$n), function(x) { 
                         print(x)
                         heatmapGenes(markers.df = genes$PFC, 
                                      n.genes = x, color.by =  gene.score,
                                      max.value = 5, title = "hi") 
                         } 
       )


df <- rbindlist(dfs)  %>% 
  mutate(individual.factor = as.factor(individual) %>% as.numeric) %>%
  group_by(gene) %>% 
  arrange(individual) %>% 
  mutate(individuals = paste(individual.factor, collapse = "")) %>%
  ungroup()

# df %>% ungroup %>% select(individual, individual.factor) %>% distinct
#df$individuals %>% unique %>% sort
#factor(levels = unique(df$individuals %>% as.character %>% sort))

df %<>% 
  arrange(individuals) %>%
  mutate(# gene = factor(as.character(gene), levels = unique(as.character(gene))),
         individuals = factor(individuals, levels = df$individuals %>% unique %>% sort),
         gene.score = gene.score.x) %>%
  select(gene, n, individual , individual.factor, individuals, -gene.score.y, everything())

df$individuals %>% levels
# Biuld plot

# df %>% select(gene, n, colors.x)

df %<>% dplyr::filter(n > 1) %>%
  mutate(gene = as_factor(as.character(gene)))

colors.x <- viridis_pal(begin = 0.2, end = 0.9)(df$n %>% max)[df %>% arrange(gene) %>% distinct(gene, n) %>% .$n]

# df %>% select(n, colors.x) %>% distinct

h <- ggplot(df) +
  geom_tile(aes(x = gene, y = individual,
                colour = gene.score.x,
                fill = gene.score.x), 
                hjust = 0.5, vjust = 0.5) +
    
   scale_fill_gradientn(aesthetics = c("colour", "fill"),
                                       na.value = "#150E37FF",
                        colors = viridis_pal(option = "magma", 
                                             begin = 0.2, end = 0.8)(100), 
                        limits = c( min(df$gene.score[! is.na(df$gene.score)]),  quantile(df$gene.score[! is.na(df$gene.score)], .95)),
                        oob = scales::squish
                        ) +
  ggtitle(df$cluster %>% unique) +
  theme_minimal() +
  theme(line = element_blank(),
        axis.text.x = element_text(size = 6, angle = 90, hjust = 1, vjust = 0.5,
                                   colour = colors.x, 
                                   margin = margin(t = -4)
                                   ),
        axis.title = element_blank(),
        aspect.ratio = .15,
        axis.ticks = element_blank()
        
        )


h
        # legend.position = "bottom")

quantile(df$gene.score, .95, na.rm = TRUE)
g$gene.score %>% summary

t <- df %>% arrange(as.character(individuals)) 
df$individuals %>% levels


# Order of genes for combined heatmaps below.
genes.order <- list("V1" = h$data$gene)
genes.order$PFC <- h$data$gene


```


```{r}
ggsave(h , filename = "../violinPlots/neocortex_DEgenes_by_celltype_area/heatmap_genescore_neurons_V1.pdf", height = 4, width = 20)
```


Bin genes into categories depending on which combination of samples they appear in
```{r}

         

# Factor correspondences
df %>%  
  ungroup %>% dplyr::select(individual, individual.factor) %>% distinct(individual, individual.factor)

# Combination counts
df %>% ungroup %>% dplyr::count(n, individuals)
```

counts <- ggplot(df) +
  geom_tile(aes(x = gene, y = 0,
                 fill = factor(n)
                 ),
                 # size = 0.1
  ) +

   theme(axis.text.x = element_text(size = 1.75, angle = 90, hjust = 1),
        axis.title = element_blank(),
        axis.text.y = element_blank()
   ) +
  
  scale_color_discrete()
        # legend.position = "bottom")

return(h)

}

levels <- data.frame(df.1$gene %>% levels,
df.2$gene %>% levels)

data.frame(df.1$gene,
df.2$gene)

lapply(list(df.1, df.2), function(x) x$gene %>% unique %>% as.character) %>% as.data.frame


patchwork::wrap_plots(h.1, h.2, byrow = TRUE)
```

ggsave(h , filename = "../violinPlots/neocortex_DEgenes_by_celltype_area/dheatmaptest.pdf", height = 4, width = 11)


  # scale_fill_viridis_c("gene.score", option = "magma", begin = 0.2, end = 0.8, limits = c(0, 15), na.value = "yellow", ) +
    # geom_label(aes(1, individual, label = n)) +
  # scale_x_discrete(limits = genes$gene) +
  # scale_y_discrete(labels = paste(genes$individual, genes$n, " ")) +

range(genes$gene.score)


markers.ct$neuron %>% str
```




## Combined: all individuals

### Neurons
```{r}
dir <- "../violinPlots/neocortex_DEgenes_by_celltype_area/markers_by_celltype_area/"
files <- paste0(dir, list.files(dir))

# Use read.delim bc 1st column is rownames (tidyr functions don't work)
markers <- lapply(files, function(x){
  
                 read.delim(x, header = FALSE) %>% 
                  set_names(c("gene", "p_val",	"avg_logFC", "pct.1", "pct.2", "p_val_adj",	"area",	"n.comps"))
            }) %>% 
            mutate(gene.ratio = pct.1 / pct.2,
                  gene.score = avg_logFC * gene.ratio
                           )
          # Name list elements according to each individual.
          set_names(value = files %>% tolower %>% str_match(pattern = ".*/(.*)_byarea") %>% .[,2]) %>%
            # https://stackoverflow.com/questions/48255351/
            # Add column with individual
            purrr::imap( ~mutate(.x, cell.type = .y))

            
```

# ------------------------------------------------------------------------------------------------------------

## Genes in common across individuals



## Top Marker genes
For each cell type, get the top marker genes for each area.
```{r}
markers %<>% lapply(function(x)
                  x %>%
                )

# write_tsv(markers, "../violinPlots/DEgenes_wholebrain_all_celltypes.txt")

# https://stackoverflow.com/a/27766224 slice_max

top.markers <- markers %>% map(dplyr::filter, n.comps >= 3) %>% 
                            map(group_by, area) %>% 
                            map(slice_max, order_by = gene.score, n = 30)



# getwrite_tsv(markers, "DEgenes_wholebrain_all_celltypes_top50.txt")
```

```{r}
s.obj <- ncx.exn

Idents(s.obj)
s.objn@meta.data %<>% set_rownames(.$cell.name)

s.obj %<>% SetIdent(value = 'cell.type')

```           


```{r}
top.markers %>% dplyr::count(area, celltype)

top.markers %<>% rbindlist

colors.area <- c("PFC"="#e72718", 
                            "motor"= "#e4840f",  
                            "somatosensory" = "#efcf60", 
                            "parietal" = "#745cda" ,
                            "temporal" = "#ff61e8",
                            "V1" = "#1b30f3" ) 

colors.area <- data.frame(area = names(colors.area), 
                          color = colors.area)
                          
                          
# for(celltype in celltypes) {
  

# Reorder areas


scales::show_col(colors.area$color, ncol = nrow(colors.area))

# https://forcats.tidyverse.org/reference/fct_relevel.html
ncx.exn@meta.data$area %<>% forcats::fct_relevel(colors.area$area)

top.markers$area %<>% forcats::fct_relevel(colors.structure$area)
  
  
stackedVlnMarkers <- function(ct, obj = s.obj) {
  
                        print(ct)
                        # print(region)
  
                        top.markers %<>% rbindlist
                        
                        lapply(colors.area$area, function(ctx.area){
                          
                            print(ctx.area) 
                          
                            genes.plot <- top.markers %>% dplyr::filter(cell.type == ct & area == ctx.area) %>% 
                                            arrange(desc(gene.score))
                            
                            print(genes.plot)
  
                            # canonical.gene <- celltypes.df %>% dplyr::filter(cell.type == ct) %>% .$marker
                            canonical.genes <- celltypes.df %>% dplyr::filter(cell.type %in% top.markers$cell.type) %>% .$marker                  
                            
                            if(nrow(genes.plot) > 0) {

                            p <- StackedVlnPlot(obj = s.obj, 
                                            features = c(canonical.genes, genes.plot$gene),
                                            title = paste0(ct, ": ", ctx.area, " markers"),
                                                        # genes.plot$gene[1: min(5, nrow(genes.plot))])
                                            idents = ct, 
                                            group.by = 'area',
                                            cols = colors.area$color)
                            
                              
                            ggsave(p, filename = paste0("../violinPlots/neocortex_DEgenes_by_celltype_area/vlnPlot_topDEgenes_neocortex_", ct, "_", ctx.area, ".pdf"), 
                                   device = "pdf", width = 10, height = 20, units = 'in')
                            
                            tw_sms(paste0("Done: ", ct))
                        
                        }
                          
                        }
                        
                        )
  

                        

}



p.violin <- 
  
  
  
lapply(celltypes.df$cell.type, stackedVlnMarkers)


tw_sms()





top.markers %>% arrange(area, desc(gene.score))
```

```{r}
ggsave(p.violin[[1]], filename = "../violinAstro.pdf", width = 10, height = 25, units = 'in')
```

```{r}
markers %>% dplyr::count(area, celltype)
```

```{r}
p_load(data.table)


top.markers <- markers[7]] %>% group_by(area) %>% 
  slice_max(order_by = gene.score, n = 20)


genes.plot <- top.markers %>% dplyr::filter(area == "GE" & celltype == 'radialglia')

top.markers %>% dplyr::count(celltype, area)


StackedVlnPlot(obj = wb.small, 
               features = c("AQP4", genes.plot$gene),
               idents = 'astrocyte', group.by = 'structure')
```

# Make featurePlots for each marker
```{r}
fPlots$astro <- top.markers$gene %>% lapply(function(x) { makePlot(gene = !!sym(x), 
                                                         cap = 4)
}
)


# https://stackoverflow.com/questions/39736655/plot-over-multiple-pages

m1 <- marrangeGrob(fPlots$astro, nrow=4, ncol=5)
names(m1) <- seq(1:length(m1))

for(i in names(m1)) {
  ggsave(plot = m1[[i]], 
         filename = paste0("../out/astrocyte_markers_", i, ".png"),
         device = "png",
         width = 45, height = 30, units = "in")
}


grid_arrange
ggarrange()
```

## Neurons
```{r}
markers[[6]] %<>% separate(cluster, c("celltype", "area")) %>% 
                      mutate(gene.ratio = pct.1 / pct.2,
                             gene.score = avg_logFC * gene.ratio)

top.markers <- list()
top.markers$neurons <- markers[[6]] %>% 
                        group_by(area) %>% 
                          slice_max(order_by = gene.score, n = 5)

fPlots$neurons <- top.markers$neurons$gene %>% 
                    lapply(function(x) { 
                      
                      makePlot(gene = !!sym(x), 
                               cap = 4)
                    }
                    )

m1 <- list()
m1$neurons <- marrangeGrob(fPlots$neurons, nrow=4, ncol=5)

names(m1$neurons) <- seq(1:length(m1$neurons))

for(i in names(m1$neurons)) {
  ggsave(plot = m1$neurons[[i]], 
         filename = paste0("../out/neuron_markers_", i, ".png"),
         device = "png",
         width = 45, height = 30, units = "in")
}
```
# ///////////////////////////////////////////////////////////////////////////////////////////////////////////

# fxn: stackedVlnPlot
```{r}

## remove the x-axis text and tick
## plot.margin to adjust the white space between each plot.
## ... pass any arguments to VlnPlot in Seurat

modify_vlnplot <- function(obj, 
                          feature, 
                          pt.size = 0,
                          ...) {
  
  p <- VlnPlot(obj, features = feature, pt.size = pt.size, ... )  + 
    
    ylab(feature) + 
    
    labs(title = NULL) +
    
    theme(legend.position = "none", 
          
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x = element_blank(),
          
          axis.title.y = element_text(size = rel(0.75), angle = 0), 
          axis.text.y = element_text(size = rel(0.75)), 
          
          plot.margin = unit(c(-2, -2, -2, -2), "cm") ) 
  
  return(p)
}

## extract the max value of the y axis
extract_max<- function(p){
  
  ymax<- max(ggplot_build(p)$layout$panel_scales_y[[1]]$range$range)
  return(ceiling(ymax))
}


## main function
StackedVlnPlot<- function(obj, 
                          features,
                          pt.size = 0,
                          title,
                          # plot.margin = unit(c(0, 0, 0, 0), "cm"),
                          ...) {
  
  plot_list <- purrr::map(features, 
                          function(x) modify_vlnplot(obj = obj,
                                                     feature = x, 
                                                     ...))
  
  # Add back x-axis title to bottom plot. patchwork is going to support this?
  plot_list[[length(plot_list)]] <- plot_list[[length(plot_list)]] +
    theme(axis.text.x = element_text(size = rel(0.75), angle = 45), 
          axis.ticks.x = element_line())
  
  # change the y-axis tick to only max value 
  ymaxs <- purrr::map_dbl(plot_list, extract_max)
  plot_list<- purrr::map2(plot_list, ymaxs, function(x,y) x + 
                            scale_y_continuous(breaks = c(y)) + 
                            expand_limits(y = y))

  p <- patchwork::wrap_plots(plotlist = plot_list, ncol = 1) +
     patchwork::plot_annotation(title = title)
  
  return(p)
}
```

# ///////////////////////////////////////////////////////////////////////////////////////////////////////////
# DotPlots

```{r}
markers.ct

# List of markers split by area.
# Rename by area.
# https://community.rstudio.com/t/using-purrr-to-rename-list-objects-in-a-list-of-tibbles-based-on-the-content-of-those-tibbles/30726/
genes %<>% set_names(., 
                    value = map(.x = ., ~first(.x$cluster))
                      )


ncx.exn@meta.data %>% str
  
) %>% filter(area %in% c("PFC", "V1")) %>% count(area, individual) %>%
  ggplot() + geom_point(aes(area, individual, size = n))


ncx.exn@meta.data$area.numeric <- as.numeric(ncx.exn@meta.data$area)
ncx.exn@meta.data$individual %<>% as.factor()
ncx.exn@meta.data$split <- ncx.exn@meta.data$area %>% factor(levels = c("PFC", "V1"))
filter(cell.type == "neuron") %>% count(cell.type.area)

Idents(ncx.exn) <- 'cell.type.area'
levels(ncx.exn)

cols <- viridis_pal(option = "magma")(25)
x <- rbind(genes$PFC, genes$V1) %>% filter(cluster %in% c("PFC", "V1") & n >= 2)
x <- genes$PFC %>% filter( n >= 2)
x %>% length

dp <- DotPlot(object = ncx.exn,
              idents = c('neuron-PFC', 'neuron-V1'),
                          features = x$gene %>% unique,
                          group.by = 'cell.type.area.indiv',
                          scale = FALSE, 
                         dot.scale = 1)
                 # split.by = 'individuak')
                # cols = cols) 
# cut.default error bc there are no GW14 PFC / GW25 V1 samples

dp <- dp + scale_color_viridis_c(option = "magma", end = 0.85)
ggsave(dp , filename = "../heatmaps_neocortex_DEgenes_celltype_x_area/dotPlot_neurons_PFC_V1.pdf", height = 5, width = 30) 


                    
p <- dp$data %>% ggplot() + geom_point(aes(x = id, y = features.plot, size = pct.exp, fill = avg.exp)) +
                                    scale_size_area(max_size = 3) 


```

# //////////////////////////////////////////////////////////////////////////////////////////////////
# Heatmaps: Gene exp acx areas / ages

```{r}
# Make column with age-area
ncx.exn@meta.data %<>% unite("group", individual, area, remove = FALSE, sep = "-")

Idents(ncx.exn) <- 'cell.type'
# Levels: cr dividing ipc neuron opc rg

# All PFC + V1 markers ( n>= 2), average expression across all age/area neuron groups.
x <- genes$PFC %>% filter( n >= 2)

dp <- DotPlot(object = ncx.exn,
              idents = 'neuron',
                       features = x$gene %>% unique,
                          group.by = 'group',
                          scale = FALSE, 
                          dot.scale = 1)


```

# fxn: heatmapCS
```{r}
df.quants <- df$avg.exp[df$avg.exp > 0] %>% quantile(c(0.1, 0.975))

# dp$data$features.plot %<>% factor(levels = genes.order$PFC %>% unique)
color.by <- quo(avg.exp)
colors <- viridis_pal(option = "magma", 
                      begin = 0.1, end = 0.8)(100)
h <- ggplot(df) +
  
  geom_tile(aes(x = gene, y = id,
                colour = log10(!!color.by),
                fill = !!color.by,
                ), 
                hjust = 0.5, vjust = 0.5) +
    
   scale_fill_gradientn(aesthetics = c("colour", "fill"),
                                       na.value = "#150E37FF",
                        colors = colors, 
                        limits = c(df.quants[1],  
                                  df.quants[2]),
                        oob = scales::squish,
                        values = scales::rescale(c(df.quants[1], 5, 15, df.quants[2]))
                        ) +
  
  # ggtitle(df$cluster %>% unique) +
  theme_minimal() +
  theme(line = element_blank(),
        axis.text.x = element_text(size = 6, angle = 90, hjust = 1, vjust = 0.5,
                                   colour = colors.x, 
                                   margin = margin(t = -4)
                                   ),
        axis.title = element_blank(),
        aspect.ratio = .5,
        axis.ticks = element_blank()
        
        )

ggsave(plot = h, "../heatmaps_neocortex_DEgenes_celltype_x_area/heatmap_geneExp_markers_PFC_1.pdf", device = "pdf", 
       height = 10, width = 40)

genes.order$PFC.V1 <- dplyr::combine(genes.order) %>% unique

ggplot(df) + geom_histogram(aes(x = avg.exp), bins = 1000) +
  geom_jitter(aes(avg.exp, 600), height = 300)
```

# Cluster genes: Reorder x axis
```{r}
df <- dp$data

g <- df %>% select(features.plot, id, avg.exp) %>% 
          pivot_wider(values_fill = -3,  
                      values_from = avg.exp,  
                      names_from = id, 
                      id_cols = features.plot)

# Hierarchical clustering
# Get dist matrix, run hcust, extract ordered labels
dd <- g %>% ungroup %>% 
    dplyr::select(-features.plot) %>% as.matrix %>% set_rownames(g$features.plot) %>%
       dist %>% hclust %>% dendro_data

# Revert to dataframe for plotting
# g <- g %>% pivot_longer(cols = 2:ncol(g), 
#                          values_to = "gene.score",  
#                           names_to = "individual") %>% 
#           dplyr::na_if(-3) %>% 
#           dplyr::filter(! is.na(gene.score))  %>%
#             left_join(markers.df,
#                       by = c("gene", "individual"))  %>%
# mutate(gene = factor(gene, levels = dd$labels$label)) %>%
# arrange(gene)

df$gene <- df$features.plot %>% factor(levels = dd$labels$label)
df$avg.exp <- round(df$avg.exp, 3)
```

## Reorder individuals: y axis
```{r}

```
