---
title: 'Constellation Plots: Neocortex Excitatory: By Ages (for reviews)'
date: 2021-03-03
output: html_document
---

```{r}
knitr::opts_knit$set(progress = TRUE, verbose = TRUE)
setwd("~/cse-phd/second-trimester/neocortex/constellation/R")
getwd()
```

```{r}
ncx.exn@meta.data %>% count(individual)
```

```{r call-buildConstellationPlot}

# Neocortex excitatory (271k) used originally to make the 'cell types by area' constellation plot.
# This object was saved when making the 'cell types by cluster' constellation plot,
# but has everything we need to make dfs.
loadData <- function(rds_path) {

    x <- read_rds(rds_path)
              
              names(x) <- c('cells_cl_df', 'cluster_colors', 'cl', 'cl_numeric', 'cl_center_df', 'reductions')
              
               x$cells_cl_df %<>%  
                                  rename(cell_type_area = "cell.type.area") %>%
                                  rename(cell_type = "cell.type") %>% 
                                  rename(cell_name = "cell.name") %>%
                 expandMetadata() %>%
                                  set_rownames(.$cell_name)
               
               # Add RG / neuron subtype from annotations table.
               x$cells_cl_df %<>% left_join(annotations %>% select(cluster_label, annotation), by = "cluster_label") %>%
                                  mutate(cell_type = as.character(cell_type),
                                         annotation = as.character(annotation),
                                         annotation = case_when(is.na(annotation) ~ cell_type,
                                                                  TRUE ~ as.character(annotation))) %>%
                                                                # TRUE ~ annotation)) %>% # IPCs
                    unite(col = 'celltype_area_2', annotation, area, remove = FALSE)
                      
                cell_names <- x$cells_cl_df$cell_name
                
                x$cl <- x$cells_cl_df$cell_type_area %>% purrr::set_names(nm = cell_names)
                x$cl_numeric <- as.numeric(as.factor(x$cl)) %>% purrr::set_names(nm = cell_names)
                x$cells_cl_df$cl_id <- x$cl_numeric
              # TODO Not converting gw16 and gw19_2 because expandMetadata needs to be fixed to include those ages.
              # WARNING: Input `stage` is `fct_collapse(...)`.Unknown levels in `f`: gw16, gw19_2
                
              return(x)
} 

getReductions <- function(object, stage_object) {
  
                    cells_cl_df <- stage_object
                    cell_names <- cells_cl_df$cell_name
                    print(cell_names)
                    reductions <- object$reductions %>% map(~.[ cell_names, ])
                    
                    # cl <- cells_cl_df$cell_type_area %>% set_names(cell_names)
                    # cl_numeric <- as.numeric(as.factor(cl)) %>% set_names(cell_names)
                    return(lst(cells_cl_df,reductions)) #cl, cl_numeric))
}

getExOnly <- function(cells) {
  
   cells$cells_cl_df %<>% 
      dfilter(cell_type %in% c('rg', 'ipc', 'neuron'))
        # mutate(annotation = str_remove_all('late| '))# get rid of 'late oRG' -> just 'oRG' | now done in main annot cleanup.
   y <- cells$cells_cl_df$cell_name
  cells$reductions %<>% map(~ .[y, ])
  return(cells)
}

prepData <- function(annotation_path,
                    cluster_colors_path,
                    out_dir,
                    object_path
                    ) {

  annotations <<- read_tsv(annotation_path)

  a <<- loadData(rds_path = object_path)
  a$cl <<- NULL
  a$cl_numeric <<- NULL
  a$cl_center_df <<- NULL

  cluster_colors <<- read_tsv(cluster_colors_path)

  a_excit <<- getExOnly(a)
  a$cells_cl_df %>% dplyr::count(cell_type, annotation, cell_type_area, celltype_area_2)

  out_dir <<- out_dir

}
# --------

x$cluster_colors %<>% rename(clade = 'celltype')
x$cells_cl_df %<>% select(-cl_id, -cluster_label) %<>% 
  rename(cluster_label = 'cell.type.area')

```

## 2. Make constellation plot.

Call `buildConstellation.Rmd`, which has all the steps required to make dfs and plot.

Takes arguments: - cells.cl.df - groups.col - colors.use - rd.dat

```{r}
'../R/buildConstellation.Rmd'

rmarkdown::render(input = "../R/buildConstellation.Rmd", 
                  params = list(cells.cl.df = x$cells_cl_df, 
                                rd.dat = x$reductions, 
                                groups.col = "cluster_label", # renamed from cell.type.area 
                                colors.use = x$cluster_colors,
                                n.pcs = 20,
                                run.knn = FALSE,
                                k = 15,
                                out.dir = '../out',
                                frac.th = 0.05,
                                cl = x$cl,
                                cl.numeric = x$cl_numeric)
)

```

## Split by stages

1.  Clean up metadata using `expandMetadata`.

This adds a colum `stage`, which groups ages into early, middle, late.

2.  Filter cells from each stage.

```{r}
early <- list()
early$cells_cl_df <- x$cells_cl_df %>% filter(stage == "early") %>% set_rownames(.$cell.name)
# 11,208

early$reductions$umap <- x$reductions$umap[early$cells_cl_df$cell.name, ]
early$reductions$pca <- x$reductions$pca[early$cells_cl_df$cell.name, ]

early$cl <- early$cells_cl_df$cluster_label %>% set_names(rownames(early$cells_cl_df))
early$cl_numeric <- as.numeric(as.factor(early$cl)) %>% set_names(rownames(early$cells_cl_df))

rmarkdown::render(input = "../R/buildConstellation.Rmd", 
                  params = list(cells.cl.df = early$cells_cl_df, 
                                rd.dat = early$reductions, 
                                groups.col = "cluster_label", # renamed from cell.type.area 
                                colors.use = x$cluster_colors,
                                n.pcs = 20,
                                run.knn = TRUE,
                                k = 15,
                                out.dir = '../out',
                                frac.th = 0.05,
                                cl = early$cl,
                                cl.numeric = early$cl_numeric)
)

```

Generalized form:

```{r}

x$cells_cl_df %>% dplyr::count(stage)

x$cells_cl_df %>% dplyr::count(individual)




split_stages %>% map(~ rmarkdown::render(input = "../R/buildConstellation.Rmd", 
                      params = list(args_list = ., 
                                
                                groups.col = "cell_type_area", # renamed from cell.type.area 
                                colors.use = x$cluster_colors,
                                n.pcs = 50,
                                run.knn = TRUE,
                                k = 15,
                                out.dir = '../out',
                                frac.th = 0.05)))






#  ---
split_stages$early %>% glimpse

split_stages %>% enframe %>% pmap(~ rmarkdown::render(input = "../R/buildConstellation.Rmd",
                                  params = list(name = ..1, args_list = ..2, 
                                  groups.col = "cell_type_area", # renamed from cell.type.area 
                                  colors.use = a$cluster_colors,
                                  max_size = 25,
                                  run.knn = TRUE,
                                  n.pcs = 20,
                                  k = 15,
                                  frac.th = 0.05,
                                  out.dir = '~/Dropbox/ARK_lab/second-trimester/2nd Trimester Areas Paper/revisions/constellation/'
                                  )))


```

With R script only (as function):
```{r}
source("../R/buildConstellation.R")

split_stages$late %>% buildConstellation(args_list = .,
                                         groups.col = "cell_type_area",
                                         # renamed from cell.type.area 
                                        colors.use = x$cluster_colors,
                                        run.knn = TRUE,
                                        k = 15,
                                        n.pcs = 20,
                                        frac.th = 0.05,
                                        out.dir = '../out'
                                      )

# All 3 stages
constellation_output <- split_stages %>% map(~ buildConstellation(args_list = .,
                                         groups.col = "cell_type_area",
                                         # renamed from cell.type.area 
                                        colors.use = x$cluster_colors,
                                        run.knn = TRUE,
                                        k = 15,
                                        n.pcs = 20,
                                        frac.th = 0.05,
                                        out.dir = '../out'
                                        ))

a$split_stages <- split_stages

write_rds(a, "./neocortex_input_for_constellation_by_stages.rds")

# TODO 2021-03-04
# - Save to Dropbox
# - Add stage to filename; remove time
# - Remove CSV save (node dodge)
# - Make scratchh hicat .R
# - Node size smaller?
# - Add oRG annotations
# - Do only between RG/Neurons/IPCs.
```

## Only between RG/IPC/excitatory neurons.
```{r 2021-03-15}

prepData(annotation_path = '../../data/tbls/Neocortex_subset1_clustermarkers_combo2_RGandNeuronannotations.csv',
         out_dir = '~/Dropbox/ARK_lab/2nd Trimester Areas Paper/revisions/Plots to Add/constellation/org_neuron_subtypes',
         cluster_colors_path = './cluster_colors.tsv',
         object_path = '../out/out_intermediate/ncx_exn_byCluster/dfs_and_objects_for_constellationPlot_ncx_exn_byCluster.rds')

split_stages_annot_ex <- a$cells_cl_df %>% split(.$stage) %>% map(~ getReductions(a, .)) %>% map(~ getExOnly(.))

source('../R/buildConstellation.R')
out_dir <- "~/Dropbox/ARK_lab/2nd Trimester Areas Paper/revisions/Plots to Add/constellation"

result <- split_stages_annot_ex %>% enframe %>% 
                  pmap(~ buildConstellation(args_list = ..2,
                                                     name = ..1,
                                                      groups.col = "cell_type_area",
                                                      colors.use = cluster_colors,
                                                      run.knn = TRUE,
                                                  node_label = 'area',
                                                      k = 15,
                                                      n.pcs = 50,
                                                      frac.th = 0.1,
                                                      out.dir = out_dir, max_size = 30,
                                            subclades = FALSE, # TODO take this out
                                            plot.parts = FALSE, dodge_value = 0.75, label.size = 2, exaggeration = 0.2))
```


## oRG, vRG, tRG

Clean up annotations.
```{r}
annotations_path <- "../../data/tbls/Neocortex_subset1_clustermarkers_combo2_RGandNeuronannotations.csv"
annotations <- read_tsv(annotations_path) %>% dfilter(!is.na(annotation)) %>%
  mutate(cluster_label = tolower(cluster_label) %>% str_remove_all('combo2_'),
         annotation = tolower(annotation) %>% str_remove_all('late|early') %>% str_replace('_', '-'),
          cell_type = str_extract(cluster_label, 'neuron|rg'))
         # annotation = paste(cell_type, annotation, sep = '_')) %>% select(cell_type, cluster_label, annotation)

write_tsv(annotations, annotations_path)
```

```{r}
annotations <- read_tsv(annotations_path)

a <- loadData(rds_path = paste0('../out/out_intermediate/ncx_exn_byCluster/',
                          'dfs_and_objects_for_constellationPlot_ncx_exn_byCluster.rds'))

a$cl <- NULL
a$cl_numeric <- NULL
a$cl_center_df <- NULL

cluster_colors <- read_tsv('../R/cluster_colors.tsv')

a$cells_cl_df %>% dplyr::count(cell_type, annotation, celltype_area_2)

split_stages_annot_ex <- 
  a$cells_cl_df %>% split(.$stage) %>% 
    map(~ getReductions(a, .)) %>% 
      map(~ getExOnly(.))

source('../R/buildConstellation.R')
out_dir <- '~/Dropbox/ARK_lab/second-trimester/2nd Trimester Areas Paper/revisions/constellation/rg_neuron_subtypes/'

result <- split_stages_annot_ex %>% enframe %>% 
                  pmap(~ buildConstellation(args_list = ..2,
                                                     name = ..1,
                                                      groups.col = "celltype_area_2",
                                                      colors.use = cluster_colors,
                                                      run.knn = TRUE,
                                                  node_label = 'area',
                                                      k = 15,
                                                      n.pcs = 50,
                                                      frac.th = 0.1,
                                                      out.dir = out_dir, max_size = 30,
                                            subclades = FALSE, # TODO take this out
                                            plot.parts = FALSE, dodge_value = 0.75, label.size = 2, exaggeration = 0.2))
```


  

```{r}
  cl.center.df <- updateClCenterDf(cl.center.df)
  
test <- result %>% map(~ updateClCenterDf(.))



cl.plots <- test %>% map(~ plot_constellation(
                              knn.cl.df = .$knn.cl.df.filter, 
                              cl.center.df = .$cl.center.df, 
                              out.dir = out_dir,
                              node.label = "cluster_label",  # name of column in cl.center.df with cluster/node names
                              exxageration = 0.2, curved = TRUE, 
                              plot.parts = FALSE, plot.hull = NULL, 
                              plot.height = 15, plot.width = 8,
                              node.dodge = TRUE, dodge_value = 1.25,
                              label.size = 1.5, max_size = 30))


 test[[1]] %>% plot_constellation(knn.cl.df = .$knn.cl.df.filter, 
                              cl.center.df = .$cl.center.df, 
                              out.dir = out_dir,
                              node.label = "cluster_label",  # name of column in cl.center.df with cluster/node names
                              exxageration = 0.2, curved = TRUE, 
                              plot.parts = FALSE, plot.hull = NULL, 
                              plot.height = 15, plot.width = 8,
                              node.dodge = TRUE, dodge_value = 1.25,
                              label.size = 1.5, max_size = 30)

test[[1]]$cl.center.df %>% as.tibble()
result[[1]]$cl.center.df %>% as.tibble()

test[[1]]$cl.center.df$cluster_size
```

## Radial glia subtypes - all stages together:

```{r}


prepData(annotation_path = '../../data/tbls/Neocortex_subset1_clustermarkers_combo2_RGandNeuronannotations.csv',
         out_dir = '~/Dropbox/ARK_lab/2nd Trimester Areas Paper/revisions/Plots to Add/constellation/org_neuron_subtypes',
         cluster_colors_path = './cluster_colors.tsv',
         object_path = '../out/out_intermediate/ncx_exn_byCluster/dfs_and_objects_for_constellationPlot_ncx_exn_byCluster.rds')

source('../R/buildConstellation.R')

result_all_stages <- 
  a_excit %>% buildConstellation(args_list = .,
                              name = 'all_stages',
                              groups.col = "celltype_area_2",
                              colors.use = cluster_colors,
                              run.knn = TRUE,
                              node_label = 'area',
                              k = 15,
                              n.pcs = 25,
                              frac.th = 0.1,
                              out.dir = out_dir, 
                              max_size = 25,     
                              subclades = FALSE, # TODO take this out
                              plot.parts = FALSE, dodge_value = 1.1, 
                              label.size = 2, exaggeration = 0.2)

# 2pm start - 35min runtime
# 5:30pm start - 

# I had not realized that there were some rg w/o a subtype annotation - those are the lone labels without a color node on the plot. 
# Should we assign those to their own ‘rg’ category, or bin them with vRGs?
# ie, were those undeterminable so we should have a pan-rg category that reflects that (edited) 


```

This lives in buildConstellation now but is unnecessary since I cleaned up the annotations.
```{r}
  updateClCenterDf <- function(x) {
    
    message('Updating cl.center.df')
    
    areas <- c('pfc', 'motor', 'somatosensory', 'parietal', 'temporal', 'v1')
    areas_rgx <- paste0('_', areas) %>% glue_collapse('|')
    
    x$cl.center.df %<>% mutate_if(.predicate = is.character, tolower) %>%
      mutate(cluster_label = str_extract(cluster_label, areas_rgx) %>% str_remove('_'),
             subclade = case_when(! .$clade == 'ipc' ~ str_remove(.$area, areas_rgx)),
      ) %>%
      unite(clade, 'clade', 'subclade') %>% 
      mutate(clade = str_remove_all(clade, '_NA| |late')) %>%
      select(-cluster_color) %>%
      left_join(cluster_colors, by = c('clade' = 'celltype'))
    
    x$knn.cl.df.filter <- filterKNN(x$knn.cl, frac.th = 0.1)
    
    print(x$cl.center.df)
    return(x)
  }