```{r}
stackedVlnMarkers <- function(ct, obj = s.obj) {
  
                        print(ct)
                        # print(region)
  
                        top.markers %<>% rbindlist
                        
                        lapply(colors.area$area, function(ctx.area) {
                          
                            print(ctx.area) 
                          
                            genes.plot <- top.markers %>% dplyr::filter(cell.type == ct & area == ctx.area) %>% 
                                            arrange(desc(gene.score))
                            
                            print(genes.plot)
  
                            # canonical.gene <- celltypes.df %>% dplyr::filter(cell.type == ct) %>% .$marker
                            canonical.genes <- celltypes.df %>% dplyr::filter(cell.type %in% top.markers$cell.type) %>% .$marker                  
                            
                            if(nrow(genes.plot) > 0) {

                            p <- StackedVlnPlot(obj = s.obj, 
                                            features = c(canonical.genes, genes.plot$gene),
                                            title = paste0(ct, ": ", ctx.area, " markers"),
                                             # genes.plot$gene[1: min(5, nrow(genes.plot))])
                                            idents = ct, 
                                            group.by = 'area',
                                            cols = colors.area$color)
                            
                              
                            ggsave(p, filename = paste0("../violinPlots/neocortex_DEgenes_by_celltype_area/vlnPlot_topDEgenes_neocortex_", ct, "_", ctx.area, ".pdf"), 
                                   device = "pdf", width = 10, height = 20, units = 'in')
                            
                            tw_sms(paste0("Done: ", ct))
                        
                            }
                        })
                    }
```

# fxn: stackedVlnPlot
```{r}
## remove the x-axis text and tick
## plot.margin to adjust the white space between each plot.
## ... pass any arguments to VlnPlot in Seurat

modify_vlnplot <- function(obj, 
                           feature, 
                           pt.size = 0,
                           ...) {
  
  p <- Seurat::VlnPlot(obj, features = feature, pt.size = pt.size, ... )  + 
    
    ylab(feature) + 
    
    labs(title = NULL) +
    
    theme(legend.position = "none", 
          
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x = element_blank(),
          
          axis.title.y = element_text(size = rel(0.75), angle = 0), 
          axis.text.y = element_text(size = rel(0.75)), 
          
          plot.margin = unit(c(-2, -2, -2, -2), "cm") ) 
  
  return(p)
}

## extract the max value of the y axis
extract_max <- function(p){
  
  ymax <- max(ggplot_build(p)$layout$panel_scales_y[[1]]$range$range)
  return(ceiling(ymax))
}


## main function
StackedVlnPlot <- function(obj, 
                          features,
                          pt.size = 0,
                          title,
                          # plot.margin = unit(c(0, 0, 0, 0), "cm"),
                          ...) {
  
  plot_list <- purrr::map(features, 
                          function(x) modify_vlnplot(obj = obj,
                                                     feature = x, 
                                                     ...))
  
  # Add back x-axis title to bottom plot. patchwork is going to support this?
  plot_list[[length(plot_list)]] <- plot_list[[length(plot_list)]] +
    theme(axis.text.x = element_text(size = rel(0.75), angle = 45), 
          axis.ticks.x = element_line())
  
  # change the y-axis tick to only max value 
  ymaxs <- purrr::map_dbl(plot_list, extract_max)
  plot_list<- purrr::map2(plot_list, ymaxs, function(x,y) x + 
                            scale_y_continuous(breaks = c(y)) + 
                            expand_limits(y = y))

  p <- patchwork::wrap_plots(plotlist = plot_list, ncol = 1) +
      patchwork::plot_annotation(title = title)
  
  return(p)
}