---
title: "Constellation Plots: Neocortex, 2nd Trimester" 
subtitle: "All Samples, All cell types"
output: html_notebook
---

```{r global_options, include=FALSE}

knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path = 'Figs/',
                      echo = FALSE, warning = FALSE, message = FALSE)

p_load("ggthemes")
```
***
 
# TO-DOs
- 8/10/20 | save wholebrain with updated @metadata as new .rds (has cluster assignments)
- 6-25-20 | Save neocortex_exn s.obj w/ counts as new .rds.

- Constellation plots for cell type / area groups.
  - Color by cell type, use label for area only.
  - Return function color.by to "clade"
  
- Plot cells as dots? To show how representative the centroid node is.

*** 

# PROCESS OUTLINE:

1. Load data.
  - Seurat object
    - @reductions: pca, umap
  - Cluster assignments table
  
2. Define which cells you're building the constellation plot with.

3. Make the dataframes needed for the constellation plot script: `constellation.Rmd`:
  1. `cells.cl.df` # s.obj@meta.data
  2. `rd.dat`: list() # s.obj@reductions
  
  3. `colors.use` : c()
  4. groups.col: ""  #character, name of column to group by


*** 

# LOAD DATA


## a) Neocortex_v3.RData
```{r load-neocortex, eval = FALSE, echo=TRUE}
# Full dataset:
# Not necessary if just plotting on already calculated PCA, UMAP, clusters.
ncx.exn <- read_rds("../data/exn_lineage/neocortex_exn_lineage_271k_noCounts.rds") # No @counts, @data.
ncx.exn@assays$RNA@counts <- read_rds("../data/exn_lineage/ncx.v3.exn_counts.rds")

# 40k subset (toy dataset)
# ncx.40k <- read_rds("constellation_plots/ncx_v3_40k.rds")

# 100k subset of excitatory neuron lineage only.
ncx.exn.100k <- read_rds("../data/exn_lineage/toy/ncx.v3.exn.sub_100k.rds")

ncx.exn@assays$RNA %>% str()

# Cluster / marker tables.
ncx.clusters <- read_delim("../tbls/83d19_Neocortex_allindividuals_combinedclusters_v1.txt",
                              "\t", escape_double = FALSE, trim_ws = TRUE)

ncx.markers <- read_delim("../tbls/8f0fc_Neocortex_subset1_clustermarkers_combo2.txt", 
                  "\t", escape_double = FALSE, trim_ws = TRUE)

```

## b) Neocortex: excitatory lineage only
```{r load-ncx-exn, eval = FALSE}
ncx.exn <- read_rds("../../data/exn_lineage/neocortex_exn_lineage_271k_noCounts.rds")
# 58G
tw_sms()
```

`ncx.exn` was built with the chunk below:

  Keep only cells with:
  - *_combo2_* `$combined.cluster.2` annotation AND
  - whose `$combined.cluster.2` annotation belongs to excitatory neuronal lineage classes.

```{r make_ncx_exn, eval=FALSE}
# Output of this chunk was saved as `/neocortex_exn_lineage_271k_noCounts.rds` (see last line.)

cells.cl.df <- ncx.clusters.exn <- 
                  ncx.clusters %>% 
                      filter(str_detect(combined.cluster.2, "combo2") ) %>% 
                        filter(str_detect(combined.cluster.2, "Neuron|CR|Dividing|RG|IPC|OPC") )


s.obj <- Neocortex %>% subsetSeurat(cells.keep = cells.cl.df$cell.name)
# 271.4K cells in excitatory lineage.

# Remove @counts slot (5GB)
s.obj@assays$RNA@counts <- matrix(c(0,0))

# Sanity check:
cells.cl.df$cell.name %>% identical(s.obj@meta.data %>% rownames)
# [TRUE]

s.obj@meta.data <- left_join(x = s.obj@meta.data %>% select(-orig.ident) %>% rownames_to_column("cell.name"),
                             y = cells.cl.df[ , c("cell.name", "combined.cluster.2")]) %>% 
                      set_rownames(.$cell.name)

# Add a celltype and celltype-area column to metadata.

ncx.exn@meta.data %<>% 
  mutate(combined.cluster.2 = str_remove(combined.cluster.2, "_combo2") %>% tolower,
         cell.type = str_replace(combined.cluster.2, "_[:digit:]+", replacement = "")) %>% 
  unite(col = "cell.type.area.indiv", cell.type, Area, Individual, sep = "-", remove = FALSE) %>%
  unite(col = "cell.type.area", cell.type, Area, sep = "-", remove = FALSE)

write_rds(ncx.exn, "../../data/exn_lineage/neocortex_exn_lineage_271k.rds")
```


## c) Wholebrain_v3.RData
  22 brain regions
  698K cells
```{r load-whole-brain, eval = FALSE, echo=TRUE}

# 1. Full dataset:
# 153Gb on hard drive
# size: 185Gb (once read in)
# Seurat object: wholebrain

obj.size(wholebrain@assays$RNA@counts)
# 11.7GB

write_rds(wholebrain@assays$RNA@scale.data, path = "../../data/wholebrain/Wholebrain_v3_scaleData.rds")
# 160 GB
tw_sms("Done: write_rds wholebrain_scaleData")

# https://rstudio.github.io/rstudioapi/reference/jobRunScript.html
jobRunScript(
  path = "../../data/Wholebrain_v3.RData",
  name = "loadWholeBrain",
  encoding = "unknown",
  workingDir = getwd(),
  importEnv = FALSE,
  exportEnv = "R_GlobalEnv"
)


# 2. Cluster assignments table:
wb.clusters <- read_delim("../data/tbls/WholeBrain_Clusteridentity_Reclustering_v2.txt", delim = "\t" )
# Clusters were collapsed from ClusterV1 (199 clusters) -> ClusterV2 (318 clusters)
# wb.clusters %>% count(Clusterv2, Clusterv1)

# ***

# Clean up table and add to @metadata:
wholebrain@meta.data %<>%
  rename_all(tolower) %>% 
    add_column(cell.name = rownames(.))

wb.clusters %<>% 
  
  rename_all(tolower) %>%
   rename(cell.name = "x1") %>% 
    mutate(cell.name = str_replace(cell.name, "Thalamus", "thalamus")) %>%
    select(cell.name:clusterv2) %>%
      right_join(wholebrain@meta.data) %>% 
         # mutate(clusterv2 = as.factor(clusterv2),
        mutate(clusterv2_id = as.numeric(clusterv2 %>% as.factor),
               cell.type.v2 = clusterv2 %>% 
                                str_extract(pattern = "RG|Dividing|Neuron|Interneuron|Astrocyte|Microglia|Vascular|Outlier|IPC|Endo|Oligo") %>% 
                                   tolower %>% str_replace("rg", "radialglia"),
                celltype.structure = paste(cell.type.v2, structure, sep = "_")
                # celltype.structure_id = as.numeric(celltype.structure %>% as.factor)) 
                # celltype_structure id (numeric) gets added after subsetting (can change depending on which groups get subsampled)
                ) %>%
                 
                  
          select(cell.name, age, individual, structure, area, area.sub, clusterv2, clusterv2_id, cell.type.v2, 
                 celltype.structure, lamina, everything() ) %>% 
      
            as.data.frame %>% set_rownames(value = .$cell.name) %>%
              .[rownames(wholebrain@meta.data), ]

write_tsv(wb.clusters, "../data/tbls/wholebrain_clusters_and_metadata.txt")           
# wb.clusters <- read_tsv("../data/tbls/wholebrain_clusters_and_metadata.txt")
# write_tsv(wholebrain@meta.data, "../data/tbls/wholebrain_metadata.txt")
wholebrain@meta.data <- read_tsv("/data/tbls/wholebrain_metadata.txt")

# wb.clusters %>% str          
# wholebrain@meta.data %>% str

# Sanity check
setdiff(wb.clusters$cell.name, wholebrain@meta.data$cell.name)
setdiff(wholebrain@meta.data$cell.name, wb.clusters$cell.name)

# 86 celltype-structure groups.
wb.clusters %>% dplyr::count(celltype.structure, celltype.structure_id)
# wb.clusters %>% dplyr::filter(is.na(structure))

wb.clusters %<>% select(-cell.name.1)
# Add wb.clusters to metadata so that subsetSeurat outputs the proper cluster info for each cell.
wholebrain@meta.data <- wb.clusters

# Small dataset for tests
# Ensure that all celltype.structure groups are represented - keep 10% of cells from each group.
wb.small <- subset(wholebrain,
                   cells= wholebrain@meta.data %>% 
                     dplyr::filter(! cell.type.v2 == "outlier") %>%
                      group_by(celltype.structure) %>% 
                        sample_frac(0.1, replace = FALSE) %>% pull(cell.name)
            )

# Check no outlier cells
wb.small@meta.data %>% dplyr::count(cell.type.v2, celltype.structure, celltype.structure_id)

write_rds(wb.small, "../data/wholebrain_v3_small68k.rds")
write_rds(wholebrain@reductions, "../data/wholebrain_v3_reductionsOnly.rds")

# wb.rd.dat <- read_rds("../data/wholebrain_v3_reductionsOnly.rds")
# wb.small <- read_rds("../data/wholebrain_v3_small68k.rds")

symdiff <- function(x,y) {
  print("x,y")
  setdiff(x,y) %>% head
  print("y, x")
  setdiff(y,x) %>% head
}


symdiff(cells$celltype.structure, wholebrain@meta.data$celltype.structure)

lapply(list(wb.small@meta.data, wholebrain@meta.data), 
       function(x) 
         dplyr::count(x, celltype.structure) %>% arrange(n)
       ) 
```

# ____________________________________________________________________________________________________________

# 1. BUILD DATAFRAMES needed for Cl plots.
     Adapted from `scrattch.hicat` 
     Calls functions in -> `buildConstellation.Rmd`.

## 1. Define cells we will use to build the constellation plot.
  
### a. Cortex: All clusters / cell types:
[All cells: 404.2K]

Keep only cells with *_combo2_* `$combined.cluster.2` annotation.
```{r eval=FALSE}

cells.cl.df <-  ncx.clusters %>% filter(str_detect(combined.cluster.2, "combo2"))

s.obj <- Neocortex %>% subsetSeurat(cells.keep = cells.cl.df$cell.name)
# 348K cells

# Sanity check
cells.cl.df$cell.name %>% identical(s.obj@meta.data %>% rownames)
[TRUE]
```


```{r}
#-------------------------------------------------------------------------------------------------------------
# Needed for dataframes:

# 1. colors for each group
# 2. cells.cl.df: df of all cells and their group/cluster assignments
# 3. dimensionality reductions: pca and umap
# 4. cl: vector of each cell's cluster/group assignment (string), where cell names are vector names.
# 5. cl numeric: vector of each cell's cluster/group assignment (numeric), where cell names are vector names.


# Fxn parameters


# Which column in @metadata to use for cluster/group assignments.
cluster.label.var <- "combined.cluster.2"

s.obj <- ncx.exn
colors.df <- celltypes.df

reductions <- s.obj@reductions %<>% 
                lapply( 
                   
                  function(x) {
                     
                      x@cell.embeddings %>% 
                        set_rownames( tolower(rownames(.) )
                        )
                   } )

cells.cl.df <- s.obj@meta.data %>% 
                rename(cluster_label = cluster.label.var) %>%
                add_column(cl_id =  .$cluster_label %>% as.factor %>% as.numeric,
                           .after = "cluster_label") %>%  
                mutate(cell.type = tolower(cell.type),
                      cell.name = tolower(cell.name)) %>% 
                      as.data.frame %>% set_rownames(.$cell.name)

# Sanity check; cl_id must be sequential (no gaps).
cells.cl.df %>% dplyr::count(cluster_label, cl_id)

# Make `cl` named vectors with both numeric and char cluster_ids.
cl <- cells.cl.df$cluster_label %>% set_names(cells.cl.df$cell.name)
cl.numeric <- cells.cl.df$cl_id %>% set_names(cells.cl.df$cell.name)

cl.df <- cells.cl.df %>% 
  dplyr::count(cluster_label, cl_id, cell.type, name = "cluster_size") %>%
            left_join(colors.df, by = c("cell.type" = "celltype")) %>%
               set_rownames(.$cluster_label)

# Sanity check; cl_id must be sequential.
cl.df %>% dplyr::count(cluster_label, cl_id)

myParams <-  list(
  
    "cells.cl.df" = cells.cl.df,

    "colors.use" = cells.cl.df,

    "cl" = cl,
    
    "cl.numeric" = cl.numeric,
    
    "n.pcs" = 10,
    
    "k" = 15,
    
    "frac.th" = 0.1,

    "cl.df" = cl.df,

    "rd.dat" = reductions

    )

rmarkdown::render("buildConstellation.Rmd", 
                  params = myParams
           )

tw_sms()
```

# ............................................................................................................

### b. Whole brain (all regions)

1. Colors
```{r colors}
colors.clade <- data.frame(# clade = unique(cl.df$clade),
                        cluster_color = c("cr"= "darkgrey", 
                                          "dividing" = "darkkhaki", 
                                          "neuron" = "deepskyblue", 
                                          "interneuron" = "deeppink2", 
                                          "ipc" = "brown4", 
                                          "microglia" = "darkorchid1",
                                          "oligo" = "cadetblue", 
                                          "outlier" = "darkslateblue", 
                                          "radialglia" = "darkorange", 
                                          "endo" = "blanchedalmond",
                                          "astrocyte" = "purple")

                        ) %>% rownames_to_column("celltype")




"caudate"         , "cerebellum"      , "CGE"            ,  "cingulate"      ,  "claustrum"     ,  "hippocampus"     
"hypothalamus"    , "insula"          , "LGE"            ,  "MGE"            ,  "midbrain"      ,  "motor"           
"nucleusaccumbens", "parietal"        , "PFC"            ,  "preoptic"       ,  "putamen"       ,  "somatosensory"   
"striatum"        , "temporal"        , "thalamus"       ,  "V1"              
,

```


2. `cl.df`: table of clusters (or cell-type/area groups)

# By celltype-region
```{r}
# rm(cl.center.df, 
# knn, knn.cl.cl.counts, knn.cell.cl.counts, knn.dist, knn.result, pred.result, rd.cl.center, rd.dat, reductions, cl.knn.dist.mean, cl.knn.dist.sd, cl.knn.dist.th, groups.col, from.size, by, i, k, knn.dist.th, m, n, pred.prob, select.cells, q, to.size, total, nn.cl.df, outlier, knn.cl.df, nn.cl.df, s.obj)

dfilter <- dplyr::filter

base.dir <- "../data/"
wb.meta <- read_tsv(paste0(base.dir, "tbls/wholebrain_metadata.txt"))
wb.reductions <- read_rds(paste0(base.dir, "wholebrain_v3_reductionsOnly.rds")) 

wb.reductions %<>% lapply( 
                   function(x) {
                     
                      x@cell.embeddings %<>% 
                        set_rownames( tolower(rownames(.) )
                        )
                     return(x)
                   } )

                   
# Remove outliers
wholebrain %<>% subset(cells = wholebrain@meta.data %>% 
                                  dplyr::filter(! cell.type.v2 == "outlier") %>% 
                                    pull(cell.name))

df.meta <- wb.meta %>% 
            dfilter(! cell.type.v2 == "outlier") %>%
              # sample_frac(size = 0.1) %>% 
  mutate(cell.name = tolower(cell.name)
              )

reductions <- 
            lapply(wb.reductions, 
                   function(x) {
                      x@cell.embeddings %<>% set_rownames(tolower(rownames(.)))
                      str(x)
                      x@cell.embeddings %<>% .[df.meta$cell.name, ]
                      return(x@cell.embeddings)
    }
    )


#-------------------------------------------------------------------------------------------------------------
# Needed for dataframes:

# 1. colors for each group
# 2. cells.cl.df: df of all cells and their group/cluster assignments
# 3. dimensionality reductions: pca and umap
# 4. cl: vector of each cell's cluster/group assignment (string), where cell names are vector names.
# 5. cl numeric: vector of each cell's cluster/group assignment (numeric), where cell names are vector names.


# Fxn parameters

# Which column in @metadata to use for cluster/group assignments.
cluster_label_col <- "celltype.structure"

cells.cl.df <- df.meta %>% 
                rename(cluster_label = cluster_label_col) %>%
                add_column(cl_id =  .$cluster_label %>% as.factor %>% as.numeric,
                           .after = "cluster_label") %>%                
  as.data.frame %>% set_rownames(.$cell.name)

# Sanity check; cl_id must be sequential (no gaps).
cells.cl.df %>% dplyr::count(cluster_label, cl_id)

# Make `cl` named vectors with both numeric and char cluster_ids.
cl <- cells.cl.df$cluster_label %>% set_names(cells.cl.df$cell.name)
cl.numeric <- cells.cl.df$cl_id %>% set_names(cells.cl.df$cell.name)

cl.df <- cells.cl.df %>% 
  dplyr::count(cluster_label, cl_id, cell.type.v2, name = "cluster_size") %>%
            left_join(colors.clade, by = c("cell.type.v2" = "celltype")) %>%
               set_rownames(.$cluster_label)

# Sanity check; cl_id must be sequential.
cl.df %>% dplyr::count(cluster_label, cl_id)

myParams <-  list(
  
    "cells.cl.df" = cells.cl.df,

    "colors.use" = colors.clade,

    "cl" = cl,
    
    "cl.numeric" = cl.numeric,
    
    "n.pcs" = 10,
    
    "k" = 15,
    
    "frac.th" = 0.1,

    "cl.df" = cl.df,

    "rd.dat" = reductions

    )

rmarkdown::render("buildConstellation.Rmd", 
                  params = myParams
           )

tw_sms()

dplyr::filter(s.obj@meta.data, is.na(cell.type.v2)) %>% dplyr::count(individual, structure)
cl.df %>% dplyr::count(cluster_label, cl_id)
```

# By clusterv2
```{r}

s.obj <- wb.small

cl <- wb.clusters$clusterv2 %>% set_names(wb.clusters$cell.name)
cl.numeric <- wb.clusters$clusterv2 %>% set_names(wb.clusters$cell.name)
cl.df <- wb.clusters %>% 
  dplyr::count(clusterv2, cl_id, cell.type.v2, name = "cluster_size") %>%
            left_join(colors.clade, by = c("cell.type.v2" = "group")) %>%
              rename(cluster_label = clusterv2) %>% set_rownames(.$cluster_label)

myParams <-  list(
  
    "cells.cl.df" = s.obj@meta.data,

    "colors.use" = colors.clade,

    "cl" = cl[s.obj@meta.data$cell.name],

    "cl.df" = cl.df[cl.df$cluster_label %in% cl[rownames(s.obj@meta.data)], ],

    "rd.dat" = list(umap = s.obj@reductions$umap@cell.embeddings,
                 pca = s.obj@reductions$pca@cell.embeddings)

    )

rmarkdown::render("buildConstellation.Rmd", 
                  params = myParams
           )

rm(cl.center.df, cl.df.discrepancies, knn, knn.cl.cl.counts, knn.cell.cl.counts, knn.dist, knn.result, pred.result, rd.cl.center, rd.dat, reductions, cl.knn.dist.mean, cl.knn.dist.sd, cl.knn.dist.th, groups.col, from.size, by, i, k, knn.dist.th, m, n, pred.prob, select.cells, q, to.size, total, nn.cl.df, outlier, knn.cl.df, nn.cl.df, s.obj)

```


```{r}


# For coloring by area-celltype

# cl.df$area <- str_split_fixed(cl.df$cluster_label, "-", 2)[ ,2] %>% tolower 

# cl.df$clade <- str_split_fixed(cl.df$cluster_label, "-", 2)[ ,1] %>% tolower 
# Add clade_id, clade_color to cl.df
# cl.df <- cl.df %>% left_join(clade.cols)



# %>%  rename(cluster_color = "colors.use")
# rm(group.cols)

# cells.cl.df: Add cluster_id column from cl.df; remove unused columns. 
# cells.cl.df <- left_join(cells.cl.df,
                         # %>% select(cell.name, groups.col, combined.cluster.2),
                         # cl.df,
                         # %>% select(cluster_label, cluster_id), 
                         by = c(groups.col = "cluster_label")
) %>% mutate(cluster_id = as.factor(cluster_id))
```
# ____________________________________________________________________________________________________________

# By cell type-area groups.

## 1. Add a cell.type-area column to metadata.
```{r }
ncx.exn@meta.data %<>% 
  mutate(combined.cluster.2 = str_remove(combined.cluster.2, "_combo2"),
         cell.type = str_replace(combined.cluster.2, "_[:digit:]+", replacement = "")) %>% 
  unite(col = "cell.type.area.indiv", cell.type, Area, Individual, sep = "-", remove = FALSE) %>%
  unite(col = "cell.type.area", cell.type, Area, sep = "-", remove = FALSE)
```

Make sure there are 271k cells in metadata.
```{r}
ncx.exn@meta.data %>% dplyr::count(cell.type)
```

```{r}
ncx.exn@meta.data %>% dplyr::count(cell.type.area)
```

## 2. Make constellation plot.

Call `buildConstellation.Rmd`, which has all the steps required to make dfs and plot.
Takes arguments: `cells.cl.df`, `groups.col`, `colors.use` and `rd.dat`.

```{r call-buildCLplot}

reductions <- list(umap = ncx.exn@reductions$umap@cell.embeddings,
                   pca  = ncx.exn@reductions$pca@cell.embeddings)

# Colors for areas:
# Add clade_id, clade_color to cl.df
colors.area <- data.frame(
                 c("motor" = "#007EB8", #command-blue
                "parietal" = "#0CD4AD", #caribbean-green
                 "pfc" = "#FF0558", #folly-red
                 "somatosensory" = "#FDD913", # yellow
                 "v1" = "#FE8499", #tulip-pink
                 "temporal" = "#C487F6" #pale-violet
                )  
              ) %>% rownames_to_column("group")


# Colors for cell types:
colors.clade <- data.frame(# clade = unique(cl.df$clade),
                        cluster_color = c("cr"= "darkgrey", 
                                  "dividing" = "darkkhaki", 
                                  "neuron" = "deepskyblue", 
                                  "inteneuron" = "deeppink2", 
                                  "ipc" = "brown4", 
                                  "microglia" = "darkorchid1",
                                  "opc" = "cadetblue", 
                                  "other" = "darkslateblue", 
                                  "rg" = "darkorange", 
                                  "vascular" = "blanchedalmond")
            ) %>% rownames_to_column("group")


rmarkdown::render("buildConstellation.Rmd", 
                  params = list(cells.cl.df = ncx.exn@meta.data,
                                rd.dat = reductions,
                                 groups.col = "cell.type.area",
                                 colors.use = colors.clade
                                 )
           )
```


****************************************************************************************************

# 3. DE genes
Find DE genes between connected clusters/nodes.

```{r diffexp}

# 1
# Dataframe w/ the proportion of k(15) nearest neighbors in each cluster for every cell.
# $nn.cl: nn = number of cells in cluster cl.
nn.cl.df <- knn.cl$pred.result$pred.prob %>% as.data.frame

# Possibly move as.factor() to top, before making all DFs.
# cl.df$cluster_id %<>% as.factor()
# cl.df %<>% mutate(cluster_label = str_remove(cluster_label, "_combo2"))

# 2
cells.cl.df %<>% rename(cluster_label = "cell.type.area")
               # rename(cluster_label = "combined.cluster.2") %>% 
               # mutate(cluster_label = str_remove(cluster_label, "_combo2"))



# Add column with cells' own cluster assignment from `cells.cl.df`.
nn.cl.df %<>% left_join(cells.cl.df,
                        by = c("query" = "cell.name")
                        )
# Add cluster_label corresponding to nn.cl.
# cl.df %<>% mutate(cluster_id = as.numeric(cluster_id))
nn.cl.df %<>% mutate(nn.cl = as.numeric(nn.cl))

nn.cl.df %<>% left_join(cl.df %>% select(cluster_label, cluster_id),
                        by = c("nn.cl" = "cluster_id"))

nn.cl.df %<>% select(query, 
                     cluster_id_self = "cluster_id", 
                    cluster_label_self = "cell.type.area",
                    cluster_id_nn = "nn.cl",
                    cluster_label_nn = "cluster_label.y",
                    freq = "Freq")
# nn.cl.df %<>% select(query, combined.cluster.2, cluster_id, nn.cl, )
```



```{r}
x <- filter(knn.cl$knn.cl.df, frac >= 0.1 & cl.from != cl.to) %>% arrange(cl.from)
```


```{r}
knn.cl$knn.cl.cl.counts %>% head
```

For each cell in cluster A, how many of its k nearest neighbors (out of k = 20) are in cluster B ?
```{r}
x <- filter(nn.cl.df, cluster_label_self == "neuron-PFC" & cluster_label_nn == "Neuron-motor")
# 12,201 cells in Neuron_8

x %>% filter(freq > 0) %>% 
  ggplot() + geom_density(aes(freq))
```

## Find `cluster_a` cells with nearest neighbors in `cluster_b`.

```{r}
# cell.cl.counts: matrix of all cells and their counts of nearest neighbors in each cluster.
cell.cl.counts <- knn.cl$knn.cell.cl.counts %>% as.data.frame.matrix %>% rownames_to_column("cell.name")

cell.cl.counts <- left_join(cells.cl.df, cell.cl.counts,  by = "cell.name") %>%
  select(-cell.type) %>% mutate(combined.cluster.2 = str_remove(combined.cluster.2, "_combo2")) %>%
  rename(cluster_label = "combined.cluster.2")


names(cell.cl.counts)[4:ncol(cell.cl.counts)] %<>% paste0("cl_", .)
```
➡️
# Find cells in `cluster_a` with nearest neighbors in `cluster_b`
```{r echo=TRUE}
cluster_a <- "IPC_6"
cluster_b <- "IPC_3"

cluster_b_col <- cell.cl.counts %>% filter(cluster_label == cluster_b) %>% 
                .$cluster_id %>% as.numeric %>% unique %>% paste0("cl_", .)


x <- cell.cl.counts %>% filter(cluster_label == cluster_a) %>% replace(is.na(.), 0)

nn.counts <- x[[cluster_b_col]]

df <- list(
n.cells.clust.a = length(x$cell.name),
with.neighbors = sum(nn.counts > 0),
no.neighbors = sum(nn.counts == 0),
median.nnCounts = median(nn.counts[nn.counts > 0])
)

as.tibble(df)


```

## Find median.
Distribution of neighbor counts in cluster_b for cells in a given cluster_a.
Use this to find the point at which cells will be split into comparison groups.
```{r median}
  ggplot(x) + 
    ggtitle(paste(cluster_a, "cells \n n/k=15 nearest neighbors in", cluster_b)) +
    
    geom_bar(aes(x = get(cluster_b_col))) +
    geom_vline(xintercept = median.nnCounts, colour = "red") +
    annotate("text", x = median.nnCounts + 1, y = quantile(1: n.cells, .07),
             label = paste0("median = ", median.nnCounts)
  ) + 
    xlab(paste("# of", cluster_b, "neighbors")) +
    ylab("") +
    theme_minimal() 
```

## FindMarkers 
Compare cells above and below the median count of `cluster_b` neighbors.
```{r findmarkers, echo=TRUE}

cells <- list()
cells$above.median <- x %>% filter(get(cluster_b_col) > median.nnCounts) %>% pull(cell.name)
cells$below.median <- x %>% filter(get(cluster_b_col) < median.nnCounts)  %>% pull(cell.name)
cells$other <- setdiff(s.obj@meta.data$cell.name, flatten_chr(cells))


s.obj <- SetIdent(object = s.obj, cells = cells$above.median, value = 'nn_ct_above_med') %>% 
                SetIdent(cells = cells$below.median, value = 'nn_ct_below_med') %>%
                SetIdent(cells = cells$other, value = 'other')

cells <- cells[1:2]

s.obj@active.ident %>% table %>% as.data.frame()
# nn_ct_below_med	7594			
# nn_ct_above_med	2557	

# Make sure @counts slot is present for slot = "scale.data" to work.
markers <- FindMarkers(s.obj, slot = "scale.data",
                          # features = s.obj@assays$RNA@var.features,
                                  ident.1 = "nn_ct_above_med", 
                                  ident.2 = "nn_ct_below_med", 
                                  logfc.threshold = 0)

```

## Clean DE genes table.
Calculate enrichment ratio, filter genes w. adj p-value < 0.05, sort table.
Positive values indicate that the feature is more highly expressed in the first group.
```{r filter-markers, echo=TRUE}
markers %<>% rownames_to_column("gene") %>%
  mutate(enrich.ratio = pct.1 / pct.2,
         gene.score = avg_diff * enrich.ratio) # what does this do to down genes?
         
markers$filter <- 
         # , across(.cols = where(is.numeric), .fns = round, digits = 4)
  # ) %>%
  filter(p_val_adj <= 0.01) %>% 
  filter(pct.1 >= 0.5 | pct.2 >= 0.5) %>%
  select(gene, pct.1, pct.2, enrich.ratio, avg_diff, gene.score, avg_diff, p_val_adj) %>%
  arrange(desc(gene.score))

markers.tmp[2:6] %<>% round(5)

write_tsv(markers.tmp, path =  paste0("../out/DEgenes_", cluster_a, "_vs_", cluster_b, "scale.data.tsv"))

```

```{r reactable-table}
reactable(markers.tmp, defaultPageSize = 100,
          showSortable = TRUE, resizable = TRUE, highlight = TRUE, filterable = TRUE, minRows = 10,
          style = list(fontFamily = "Work Sans, sans-serif", fontSize = "12px")
  )
```

```{r}
# saveWidget(markers.tmp, file = )

# Same genes in both comparisons (sanity check)
xtab_set <- function(A,B){
              both    <-  union(A,B)
              inA     <-  both %in% 
              inB     <-  both %in% B
              return(table(inA,inB))
            }
# xtab_set(markers$nn_above_med$geme, markers$nn_below_med$gene)
```

****************************************************************************************************

# Heatmmap of DE genes

Requires: 
- Seurat object
- cells.to.plot (chr vector)
- markers.to.plot (chr vector)

```{r build-heatmap}
markers <- read_tsv("../out/DEgenes_IPC_6_vs_IPC_3scale.data.tsv")
```
# Filter markers to include in heatmap

By:
- % cells expressing
- enrichment ratio
```{r}
markers.to.plot <- markers %>% 
  filter(pct.1 >= 0.4 | pct.2 >= 0.4) %>%
    filter(enrich.ratio >= 1.25 | enrich.ratio <= 0.75) %>%
  arrange(desc(enrich.ratio)) %>% pull(gene)

cells.to.plot <- flatten_chr(cells)

# Get 1st and 99th percentiles to choose the limits for color scale. (clip outliers)
(exp.limits <- s.obj@assays$RNA@scale.data[markers.to.plot, cells.to.plot] %>% as.numeric %>%
                  quantile(c(0, 0.01, 0.05, 0.1, 0.5, 0.9, 0.95, 0.99, 1)))

message(paste("Building heatmap for", length(markers.to.plot), "genes and", length(cells.to.plot), "cells."))


heatmap <- 
  Seurat::DoHeatmap(s.obj, 
            cells = cells.to.plot,
            features =  markers.to.plot,
            disp.min = exp.limits["1%"],
            disp.max = exp.limits["99%"],
            angle = 0,
            slot = "scale.data"
          # slim.col.label = TRUE,
          # remove.key = TRUE
  ) +
  theme(legend.text = element_text(size = 8),
        legend.position = "bottom",
        #legend.position = "none",
        text = element_text(size = 8),
        aspect.ratio = c(2,1)
  ) +
    scale_fill_viridis(end = 1, na.value = 'white', option = "magma", discrete = FALSE)

ggsave(heatmap, filename = paste0("../out/DEgenes_", cluster_a, "_vs_", cluster_b, "scale.data_heatmap.pdf") )

saveplot(filename = paste0("../out/DEgenes_", cluster_a, "_vs_", cluster_b, "scale.data_heatmap.pdf"),heatmap )
#, width = 11, height = 8, units = "in"
p <- ggplotly(heatmap, tooltip = "Feature", width = 1000, height = 800) %>% 
      layout(legend = list(yanchor = 'bottom', orientation = 'h')) %>% partial_bundle()


# Remove group key.
# plotly_json(p)
p$x$data[c(2,3)] <- NULL

# saveWidget("~/carmensandoval.github.io/arklab/2nd-tri/clustering/DEgenes_", cluster_a, "_vs_", cluster_b, "scale.data.html", selfcontained = TRUE)

saveWidget(p, paste0("../out/DEgenes_", cluster_a, "_vs_", cluster_b, "scale.data_heatmap.html"), selfcontained = TRUE)
```

***



# Matrix: Individual/Area cell counts
```{r}
count(cells.cl.df, Individual, Area) %>% 
  pivot_wider(names_from = Area, values_from = n)

x <- count(cells.cl.df, Individual, Area)
x$Area <- factor(x$Area, levels = c("PFC", "motor", "somatosensory", "parietal", "temporal", "V1"))
```

## fxn: plotCellCounts
```{r}

plotCellCounts <- function(counts.matrix, title) {
  
  p <- ggplot(counts.matrix) + 
  geom_point(aes(x = Area, y = Individual, size = n, color = Area)) + 
  geom_label(aes(x = Area, y = Individual, size = n, label = round(n/1000, 1), fill = Area), 
             nudge_x = 0, nudge_y = 0, color = "white", label.r = unit(0.3, "lines")) +
  theme_minimal() + 
  scale_size_area(max_size = 8) +
  # scale_radius(range = c(3,6)) +
  scale_color_viridis_d(option = "C", end = 0.85) + 
  scale_fill_viridis_d(option = "C", end = 0.85) + 
  guides(color = FALSE,
         size = FALSE,
         fill = FALSE) +
  theme(axis.text.x = element_text(angle = 35),
        text = element_text(size=12),
          axis.title.x = element_blank()
        ) +
    ggtitle(label = title,
            subtitle = "cells per sample / area (x 1K)")
  
  return(p)
}

```

```{r}
p %>% ggplotly(tooltip = "n") %>% 
  layout(showlegend = FALSE, width = 500, height = 500)

p %>% saveWidget(file = "cellCounts_indiv_area.html")
```

```{r}
cells.cl.df %<>% mutate(Area = factor(cells.cl.df$Area,
                                      levels = c("PFC", "motor", "somatosensory", "parietal", "temporal", "V1")
                                      )
                  )
```

Neurons only:
```{r}
x <- cells.cl.df %>%
  filter(clade == "neuron") %>%
        count(Individual, Area)

x$Area <- factor(x$Area, levels = c("PFC", "motor", "somatosensory", "parietal", "temporal", "V1"))

(p.neurons <- plotCellCounts(counts.matrix = x, title = "Neurons"))
# ggsave(filename = "../out/countsPlot_neurons.pdf", plot = p.neurons, width = 7, height = 6, units = "in")
```

Dividing cells:
```{r}
x <- cells.cl.df %>%
  filter(clade %in% c("rg", "dividing")) %>%
        count(Individual, Area)

x$Area <- factor(x$Area, levels = c("PFC", "motor", "somatosensory", "parietal", "temporal", "V1")
                 )

(p.div <- plotCellCounts(counts.matrix = x, title = "Dividing cells"))
# ggsave(filename = "../out/countsPlot_dividing.pdf", width = 7, height = 6, units = "in")
```

All cells:
```{r}
x <- cells.cl.df %>%
        count(Individual, Area)

x$Area <- factor(x$Area, levels = c("PFC", "motor", "somatosensory", "parietal", "temporal", "V1")
                 )

(p.all <- plotCellCounts(counts.matrix = x, title = "All cells"))
# ggsave(filename = "../out/countsPlot_allCells.pdf", width = 7, height = 6, units = "in" )
```

```{r}
plot_grid(p.all, p.div, p.neurons, nrow = 3, labels = c("", "", ""),
          rel_heights = c(1, 1, 1))

ggsave(filename = "../out/countsPlots.pdf", width = 7, height = 16, units = "in" )
```


```{r}
datapasta::tribble_paste(x)
data.frame(
  stringsAsFactors = FALSE,
        Individual = c("GW14","GW14","GW14","GW17",
                       "GW17","GW17","GW17","GW17","GW18","GW18","GW18",
                       "GW18","GW18","GW18_2","GW18_2","GW18_2","GW18_2",
                       "GW18_2","GW19","GW19","GW19","GW19","GW19","GW19",
                       "GW20","GW20","GW20","GW20","GW20_31","GW20_31",
                       "GW20_31","GW20_31","GW20_34","GW20_34","GW20_34",
                       "GW20_34","GW22","GW22","GW22","GW22","GW22","GW22T",
                       "GW22T","GW22T","GW22T","GW25","GW25","GW25","GW25",
                       "GW25"),
                 n = c(162L,1687L,996L,340L,457L,
                       64L,4L,503L,378L,42L,500L,563L,827L,2524L,1317L,
                       4842L,4469L,4368L,637L,2115L,3299L,1386L,3696L,
                       956L,762L,3225L,1122L,1229L,1063L,271L,1092L,1436L,
                       1693L,4L,1241L,802L,1061L,1244L,2829L,904L,631L,
                       5L,774L,999L,1234L,2883L,842L,1204L,7413L,222L),
              Area = as.factor(c("motor",
                                 "somatosensory","V1","PFC","motor",
                                 "somatosensory","parietal","V1","PFC","motor",
                                 "parietal","temporal","V1","motor",
                                 "somatosensory","parietal","temporal","V1","PFC","motor",
                                 "somatosensory","parietal","temporal","V1",
                                 "PFC","motor","somatosensory","V1","PFC",
                                 "parietal","temporal","V1","PFC","motor",
                                 "parietal","V1","PFC","motor",
                                 "somatosensory","parietal","V1","PFC","motor",
                                 "somatosensory","parietal","PFC","motor",
                                 "somatosensory","parietal","temporal"))
)

x <- data.frame(Individual = c("GW14",
                                 "GW14","GW14","GW17","GW17","GW17","GW17",
                                 "GW17","GW18","GW18","GW18","GW18","GW18",
                                 "GW18_2","GW18_2","GW18_2","GW18_2","GW18_2",
                                 "GW19","GW19","GW19","GW19","GW19",
                                 "GW19","GW20","GW20","GW20","GW20","GW20_31",
                                 "GW20_31","GW20_31","GW20_31","GW20_34",
                                 "GW20_34","GW20_34","GW20_34","GW22","GW22",
                                 "GW22","GW22","GW22","GW22T","GW22T","GW22T",
                                 "GW22T","GW25","GW25","GW25","GW25",
                                 "GW25"),
                           n = c(312L,3432L,
                                 635L,557L,717L,79L,5L,583L,9687L,
                                 12947L,1332L,1355L,7180L,3171L,2305L,3162L,
                                 6276L,25L,4748L,2383L,5L,2647L,1374L,2719L,
                                 2939L,20L,2184L,5441L,4903L,2365L,3150L,
                                 1844L,2863L,15L,5882L,3872L,3544L,1503L,
                                 708L,821L,846L,24L,1367L,1538L,719L,
                                 17223L,716L,1157L,10400L,767L),
                        Area = as.factor(c("motor","somatosensory",
                                           "V1","PFC","motor","somatosensory",
                                           "parietal","V1","PFC","motor",
                                           "parietal","temporal","V1","motor",
                                           "somatosensory","parietal",
                                           "temporal","V1","PFC","motor",
                                           "somatosensory","parietal","temporal","V1",
                                           "PFC","motor","somatosensory","V1",
                                           "PFC","parietal","temporal","V1",
                                           "PFC","motor","parietal","V1",
                                           "PFC","motor","somatosensory",
                                           "parietal","V1","PFC","motor",
                                           "somatosensory","parietal","PFC","motor",
                                           "somatosensory","parietal","temporal")
                                )
)

x <- data.frame(
  stringsAsFactors = FALSE,
        Individual = c("GW14","GW14","GW14","GW17",
                       "GW17","GW17","GW17","GW17","GW18","GW18","GW18",
                       "GW18","GW18","GW18_2","GW18_2","GW18_2","GW18_2",
                       "GW18_2","GW19","GW19","GW19","GW19","GW19","GW19",
                       "GW20","GW20","GW20","GW20","GW20_31","GW20_31",
                       "GW20_31","GW20_31","GW20_34","GW20_34","GW20_34",
                       "GW20_34","GW22","GW22","GW22","GW22","GW22","GW22T",
                       "GW22T","GW22T","GW22T","GW25","GW25","GW25","GW25",
                       "GW25"),
                 n = c(312L,3432L,635L,557L,717L,
                       79L,5L,583L,9687L,12947L,1332L,1355L,7180L,3171L,
                       2305L,3162L,6276L,25L,4748L,2383L,5L,2647L,1374L,
                       2719L,2939L,20L,2184L,5441L,4903L,2365L,3150L,
                       1844L,2863L,15L,5882L,3872L,3544L,1503L,708L,821L,
                       846L,24L,1367L,1538L,719L,17223L,716L,1157L,
                       10400L,767L),
              Area = as.factor(c("motor",
                                 "somatosensory","V1","PFC","motor",
                                 "somatosensory","parietal","V1","PFC","motor",
                                 "parietal","temporal","V1","motor",
                                 "somatosensory","parietal","temporal","V1","PFC","motor",
                                 "somatosensory","parietal","temporal","V1",
                                 "PFC","motor","somatosensory","V1","PFC",
                                 "parietal","temporal","V1","PFC","motor",
                                 "parietal","V1","PFC","motor",
                                 "somatosensory","parietal","V1","PFC","motor",
                                 "somatosensory","parietal","PFC","motor",
                                 "somatosensory","parietal","temporal"))
)

x <- data.frame(
  stringsAsFactors = FALSE,
           Individual = c("GW14","GW14","GW14",
                          "GW17","GW17","GW17","GW17","GW17","GW18","GW18",
                          "GW18","GW18","GW18","GW18_2","GW18_2","GW18_2",
                          "GW18_2","GW18_2","GW19","GW19","GW19","GW19",
                          "GW19","GW19","GW20","GW20","GW20","GW20",
                          "GW20_31","GW20_31","GW20_31","GW20_31","GW20_34",
                          "GW20_34","GW20_34","GW20_34","GW22","GW22","GW22",
                          "GW22","GW22","GW22T","GW22T","GW22T","GW22T",
                          "GW25","GW25","GW25","GW25","GW25"),
                 Area = c("motor","somatosensory",
                          "V1","motor","parietal","PFC","somatosensory",
                          "V1","motor","parietal","PFC","temporal","V1",
                          "motor","parietal","somatosensory","temporal","V1",
                          "motor","parietal","PFC","somatosensory","temporal",
                          "V1","motor","PFC","somatosensory","V1",
                          "parietal","PFC","temporal","V1","motor","parietal",
                          "PFC","V1","motor","parietal","PFC","somatosensory",
                          "V1","motor","parietal","PFC","somatosensory",
                          "motor","parietal","PFC","somatosensory","temporal"),
                    n = c(530L,5329L,1913L,1210L,
                          11L,935L,155L,1125L,13312L,2206L,11863L,2358L,
                          9533L,8476L,12826L,6033L,13817L,7467L,7013L,
                          5824L,7017L,4402L,6812L,5599L,3256L,3851L,4439L,
                          8216L,2844L,6371L,4658L,3598L,24L,8174L,7192L,
                          6327L,3169L,1975L,5377L,4152L,1931L,2418L,
                          2415L,32L,2946L,1923L,22743L,20608L,5956L,1033L)
   )

```


# Other functions
```{r functions}
.env$source_rmd <- function(file, local = FALSE, ...){
  options(knitr.duplicate.label = 'allow')

  tempR <- tempfile(tmpdir = ".", fileext = ".R")
  on.exit(unlink(tempR))
  knitr::purl(file, output = tempR, quiet = TRUE)

  envir <- globalenv()
  source(tempR, local = envir, ...)
}

.env$reactable <- function(...) {
  htmltools::tagList(reactable::reactable(...))
}

# source("../../../code_general/setup_R_session_CSE.R")

attach(.env)
```

